---
title: "Wigan Grassland Restoration"
author: "Kaydee S. Barker"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Set up

```{r setup, include=FALSE}

rm(list=ls()) #clear global environment/workspace
knitr::opts_chunk$set(echo = TRUE) #default display code in knit output (FALSE will hide)

```

## Load libraries (and install if needed)

```{r libraries, warning=FALSE, include=FALSE}

#Create groups of packages/libraries 

##1. For data reading and manipulation - could just do tidyverse and a few others, or make group of specific libraries for faster loading
datalib <- c("dplyr", "tidyr", "readr", "readxl", "stringr", "tibble", "reshape2", "janitor") 

##2. For data visualization (note ggplot2 is in tidyverse)
vislib <- c("ggplot2", "ggthemes", "ggfortify", "ggrepel", "plotrix", "RColorBrewer", "Cairo", "ggsci")

##3. For calculation and stats functions, including ecology-specific libraries
statlib <- c("car", "MASS", "rcompanion", "zoo", "lme4", "MuMIn", "emmeans", "ARTool", "dunn.test", "performance", "see", "corrplot", "BAT", "vegan")

#Install packages as needed - commented out, but remove # and adjust as neededas a group:
#lapply(c(datalib, vislib, statlib), install.packages)

#Load libraries
lapply(c(datalib, vislib, statlib), library, character.only = TRUE)

```

## Create code for commonly used functions, colors, settings, etc

### Summary statistics list

```{r reuse code, echo=FALSE}

#Create list of descriptive statistics to run across multiple variables with one function - will use in summarise with across
summary_stats <- list(
  n = ~length(.), #sample size
	mean = ~mean(., na.rm=TRUE), #mean
	sd = ~sd(., na.rm=TRUE), #standard deviation
	var = ~var(., na.rm=TRUE), #variance
  cv = ~sd(., na.rm = TRUE) / mean(., na.rm = TRUE), #coefficient of variation
  se = ~plotrix::std.error(., na.rm = TRUE)) #standard error

```

### Custom design settings for plots

```{r design, echo=FALSE}

#Color palettes

#color palette example using hex codes
colors <- c("#84B699", "#EF8E45", "#BF876E", "#46A9B0")
colors5 <- c("#84B699", "#EF8E45", "#BF876E", "#46A9B0", "azure")
colors9 <- c("#84B699", "#EF8E45", "#BF876E", "#46A9B0", "azure", "#A1EBCF", "#FFBC7E", "#FFCAB1", "#61E9F2")

#color palette with gradient with 20 steps
colors_ramp <- colorRampPalette(c("#84B699", "azure","#BF876E"))(25)

#Color variables

#define individual colors for plot elements
plotline_color <- "#BF876E" #define color for plot lines
fill_color <- "#84B699" #define color for fill

#Custom ggplot configurations 

#Create a custom theme for ggplot with font, legend, and background settings. Transparent allows transparent png or pdf export.
my_theme <- theme(text = element_text(#family = "Arial", 
                                      size = 12, color = "grey10"), #set universal font family, size, and color
              plot.title = element_text(size = 20, color = "black", face="bold"), #set title size, color, bold
              plot.title.position = "plot", #set title position
              legend.title = element_blank(), #remove legend title
              legend.position="right", legend.box = "vertical", #where legend appears, what orientation
              panel.background = element_rect(fill='transparent'), plot.background = element_rect(fill='transparent', color=NA), #make backgrounds transparent
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) #no gridlines

```

### Order variables

```{r labels, echo=FALSE}

#Create an order of observations within a variable to override the default order for plots
order_sites <- c("Bickershaw 1", "Bickershaw 2", "Amberswood", "Flashes")

order_management <- c("Unmanaged", "4 Years", "15 Years", "23 Years")

```

### Write functions to determine differences

```{r comparisons stats - set up function}

#pairwise comparisons

#first look at group sizes
evaluate_group_sizes <- function(data, factors, continuous_vars) {
  insufficient_groups <- list()

  for (cont_var in continuous_vars) {
    for (factor in factors) {
      filtered_data <- na.omit(data[, c(factor, cont_var)]) #remove NA values for continuous variable and factor

      group_sizes <- table(filtered_data[[factor]]) #calculate group sizes
      
      #identify groups with fewer than 2 observations
      insufficient <- group_sizes[group_sizes < 2]
      if (length(insufficient) > 0) {
        insufficient_groups[[paste(cont_var, factor, sep = "_")]] <- data.frame(
          Factor = factor,
          Continuous_Var = cont_var,
          Group = names(insufficient),
          Observations = as.numeric(insufficient)
        )
      }
    }
  }

  #combine results into a single dataframe
  if (length(insufficient_groups) > 0) {
    return(bind_rows(insufficient_groups, .id = "Combination"))
  } else {
    return(data.frame(
      Combination = character(),
      Factor = character(),
      Continuous_Var = character(),
      Group = character(),
      Observations = numeric()
    ))
  }
}

#analysis function with Bartlett's test, group size evaluation, and NA omission
differences_function <- function(data, factors, continuous_vars) {
  results_list <- list()

  for (cont_var in continuous_vars) {
    for (factor in factors) {
      # Remove NA values for the current continuous variable and factor
      filtered_data <- na.omit(data[, c(factor, cont_var)])

      # Check if all groups for the factor have at least 2 observations
      group_sizes <- table(filtered_data[[factor]])
      if (any(group_sizes < 2)) {
        # If a group has fewer than 2 observations, skip Bartlett's test and add a note
        results_list[[paste(cont_var, factor, sep = "_")]] <- data.frame(
          Comparison = NA,
          P_Value = NA,
          Continuous_Var = cont_var,
          Factor = factor,
          Test = "Skipped (Insufficient group size for Bartlett's test)",
          Significance = "Not significant"
        )
        next
      }

      #Bartlett's test for homogeneity of variances
      bartlett <- bartlett.test(as.formula(paste(cont_var, "~", factor)), data = filtered_data)

      if (bartlett$p.value > 0.05) { #for homogeneous variances: ANOVA and Tukey's HSD
        anova_res <- aov(as.formula(paste(cont_var, "~", factor)), data = filtered_data)
        tukey <- TukeyHSD(anova_res)

        #extract pairwise comparisons and P-values
        if (!is.null(tukey$comparison)) {
          sig_pairs <- data.frame(
            Comparison = rownames(tukey$comparison),
            P_Value = tukey$comparison[, "pvalue"]
          ) %>%
            mutate(
              Continuous_Var = cont_var,
              Factor = factor,
              Test = "Tukey's HSD",
              Significance = case_when(
                P_Value <= 0.001 ~ "***",
                P_Value <= 0.01 ~ "**",
                P_Value <= 0.05 ~ "*",
                TRUE ~ "Not significant"
              )
            )

          results_list[[paste(cont_var, factor, sep = "_")]] <- sig_pairs
        } else {
          #if no valid groupings are present, return an empty dataframe
          results_list[[paste(cont_var, factor, sep = "_")]] <- data.frame(
            Comparison = NA,
            P_Value = NA,
            Continuous_Var = cont_var,
            Factor = factor,
            Test = "Tukey's HSD (No significant groupings)",
            Significance = "Not significant"
          )
        }
      } else {
        #heterogeneous variances: Kruskal-Wallis and Dunn's Test
        kruskal <- kruskal.test(as.formula(paste(cont_var, "~", factor)), data = filtered_data)

        if (kruskal$p.value < 0.05) { #Dunn's Test if Kruskal-Wallis is significant
          dunn <- dunn.test(filtered_data[[cont_var]], filtered_data[[factor]], method = "bonferroni", kw = FALSE)

          #extract pairwise comparisons and P-values
          sig_pairs <- data.frame(
            Comparison = dunn$comparisons,
            P_Value = dunn$P.adjusted
          ) %>%
            filter(P_Value < 1) %>% #include all comparisons
            mutate(
              Continuous_Var = cont_var,
              Factor = factor,
              Test = "Dunn's Test",
              Significance = case_when(
                P_Value <= 0.001 ~ "***",
                P_Value <= 0.01 ~ "**",
                P_Value <= 0.05 ~ "*",
                TRUE ~ "Not significant"
              )
            )

          results_list[[paste(cont_var, factor, sep = "_")]] <- sig_pairs
        } else {
          results_list[[paste(cont_var, factor, sep = "_")]] <- data.frame(
            Comparison = NA,
            P_Value = NA,
            Continuous_Var = cont_var,
            Factor = factor,
            Test = "Kruskal-Wallis (NS)",
            Significance = "Not significant"
          )
        }
      }
    }
  }

  #combine results for all factors and continuous variables
  results_df <- bind_rows(results_list, .id = "Analysis_Combination")
  return(results_df)
}

#define what to analyze
#factors <- "Site" #predictor variable
#continuous_vars <- c("var1","var2","var3") #response variables
#data <- site_data

#evaluate group sizes
#group_size_issues <- evaluate_group_sizes(data, factors, continuous_vars) 

#run the analysis
#var_diff <- differences_function(data, factors, continuous_vars)


```


## Load data

### Soil compounds and properties
```{r load data 1, include=FALSE}

#read in site data
site_data <- read_csv("Wigan_sitechar.csv") %>%
  filter(!is.na(Site)) #remove rows with no site

#calc means and standard error for site data
site_means <- site_data %>%
  group_by(Site, Management) %>%
    dplyr::select(!Mgmt_num) %>%
    dplyr::select(!Collar_Lat) %>%
    dplyr::select(!Collar_Long) %>%
    dplyr::select(!Plot_ID) %>%
    dplyr::select(!BD_avg) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats

#pivot metals with tidyr - increase rows / decrease columns
soil_metals <- site_data %>%
	pivot_longer(
		  cols = c(10:13,15:26,28,29,31,32,34), #pick columns to pivot
		  #cols = ppm_Al:ppm_Zn, #pick columns
		  names_to = "nutrient", #name column to put original column names in
		  names_prefix = "ppm_", #remove prefix from column names
		  values_to = "ppm") %>% #name column to put values in
  dplyr::select(Site, Management, nutrient, ppm) #select columns to keep

#read in FTIR data
FTIR <- read_csv("FTIR_Sam.csv")
FTIR$Site <- as.factor(FTIR$Site)
FTIR <- FTIR[,-1]

#order variable
FTIR$Site <- factor(FTIR$Site, levels = order_sites)

#pivot
FTIR_long <- FTIR %>%
  pivot_longer(
    cols = c(2:ncol(FTIR)), #pick columns to pivot
    names_to = "wavelength", #name column to put original column names in
    values_to = "absorbance") #name column to put values in

#standardize
FTIR_standardized <- FTIR %>%
  mutate_if(is.numeric,scale)

FTIR_standardized$Site <- factor(FTIR_standardized$Site, levels = order_sites)

#numeric values only for prcomp
FTIR_standardized_num <- FTIR_standardized[,-1]


```

### Plants and fungi

```{r load data 2, include=FALSE}

#read in 13C initial harvest data
harvest <- read_csv("Stagg_harvest.csv") %>% #load
  clean_names() #clean up column names

#pivot and calculate root:shoot ratio
plant_harvest <- harvest %>%
  mutate(across(c(7:16), as.numeric)) %>% #convert columns to numeric
  dplyr::select(site, management, pot, species, functional_group, sample_type, biomass_g) %>%
  filter(biomass_g >= 0) %>% #remove non-plant data
  group_by(site, management, pot, species, functional_group, sample_type) %>%
  summarise(biomass_g = mean(biomass_g, na.rm = TRUE), .groups = 'drop') %>% #summarize duplicate entries for groups
  pivot_wider(names_from = sample_type, values_from = biomass_g) %>% #determines observation name and value columns
  mutate(root_shoot = #calculate root:shoot ratio
           case_when(root > 0 & shoot > 0 ~ root / shoot, #when values are present
                     TRUE ~ NA)) %>% #default to NA when values are not present
  mutate(harvesttime = #define harvest time
        case_when(pot == "G1" | pot == "G20" | pot == "G30" | pot == "G31" | pot == "G39" |
                  pot == "G8" | pot == "G33" | pot == "G35" | pot == "G41" | pot == "G42" |
                  pot == "G9" | pot == "G12" | pot == "G22" | pot == "GA1" | pot == "GA2" |
                  pot == "G11" | pot == "G18" | pot == "G29" | pot == "GF1" | pot == "GX"
                  ~ "T0", #time 0 beginning of incubation
                     TRUE ~ "T10mos")) #10 months incubation

#summarize biomass for plots
summary_harvest <- plant_harvest %>%
  dplyr::select(site, management, species, functional_group, shoot, root, root_shoot) %>%
  group_by(site, management, species, functional_group) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats

#summarize biomass by plant species for estimates
summary_species <- plant_harvest %>%
  dplyr::select(species, functional_group, shoot, root, root_shoot) %>%
  group_by(species, functional_group) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats

#13C calculations
nat_abun <- harvest %>%
  filter(grepl("pre", sample_type)) %>% #keep only pre-labelling data
  dplyr::select(site, sample_type, adjusted_delta, per_atom13c, per_c) %>%
  mutate(sample_type = str_replace(sample_type, "leaf pre sample", "shoot")) %>% #match naming
  mutate(sample_type = str_replace(sample_type, "root pre sample", "root")) %>% #match naming
  mutate(sample_type = str_replace(sample_type, "soil pre sample", "soil")) #match naming)

#13C df for plants only

plant13C <- harvest %>%
  filter(biomass_g >= 0) %>% #remove non-plant data
  dplyr::select(site, pot, species, functional_group, sample_type, biomass_g, adjusted_delta, per_atom13c, per_c) %>%
  mutate(total_c_mg = biomass_g * (per_c / 100) * 1000) %>% #calculate total C in mg
  mutate(excess_per13C = case_when(sample_type == "shoot" | sample_type == "root"
                     ~ per_atom13c - nat_abun$per_atom13c #calculate excess 13C
                     [match(paste(site, sample_type), 
                            paste(nat_abun$site, nat_abun$sample_type))], TRUE ~ NA)) %>% 
  mutate(excess13C_mg = case_when(sample_type == "shoot" | sample_type == "root"
                     ~ total_c_mg * (excess_per13C/100), #atom % times total C in mg
                     TRUE ~ NA)) 
  

#summarize means 13C for plots - by species
summary_plant13C_sp <- plant13C %>%
  group_by(site, sample_type, species, functional_group) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats

#summarize 13C for plots add up - by site and type only
summary_plant13C <- plant13C %>%
  dplyr::select(site, sample_type, pot, biomass_g, total_c_mg, excess13C_mg) %>%
  group_by(site, sample_type, pot) %>%
  summarise_all(funs(sum(., na.rm=TRUE))) %>% #first add up per pot
  group_by(site, sample_type) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats #then summarize per site

#add mean data to harvest data for estimates
plant_harvest_est <- plant_harvest %>%
  #replace(is.na(.), 0) %>%
  mutate(site2 =
           case_when(pot == "G20" | pot == "G31" | #higher root values
                       pot == "G37" | pot == "G7"
                     ~ "bickershaw1_high",
                     pot == "G1" | pot == "G30" | pot == "G39" | #lower root values
                       pot == "G34" | pot == "G26" | pot == "G3" 
                     ~ "bickershaw1_low",
                     pot == "G33" | pot == "G35" | pot == "G41" |
                       pot == "G23" | pot == "G24" |pot == "G40"
                     ~ "bickershaw2_high",
                     pot == "G8" | pot == "G42" |
                       pot == "G4" | pot == "G28"
                     ~ "bickershaw2_low",
                     pot == "G12" | pot == "GA1" | pot == "GA2" |
                       pot == "G5" | pot == "G6" | pot == "GA4" 
                     ~ "amberswood_high",
                     pot == "G22" | pot == "G9" |
                       pot == "G21" | pot == "GA3"
                     ~ "amberswood_low",
                     .default = site)) %>%
  mutate(root_shoot =
           case_when(root_shoot > 0 ~ root_shoot, 
                     TRUE ~ summary_species$root_shoot_mean
                     [match(species, summary_species$species)])) %>%
  mutate(root =
           case_when(root > 0 ~ root,
                     TRUE ~ root_shoot * shoot))

#summarize general roots amount by site
general_roots <- plant_harvest_est %>%
  filter(species == "general_roots") %>%
  dplyr::select(site2, root) %>%
  group_by(site2) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats

#calculate and estimate biomass per pot
total_pots <- plant_harvest_est %>%
  dplyr::select(site2, management, pot, harvesttime, root, shoot, root_shoot) %>%
  group_by(site2, management, pot, harvesttime) %>%
  summarise(root = sum(root, na.rm=TRUE), shoot = sum(shoot, na.rm=TRUE), root_shoot = mean(root_shoot, na.rm=TRUE)) %>% #summarize
  mutate(root = #add some general roots to the estimated beginning biomass for incubation
           dplyr::case_when(harvesttime == "T0" ~ root,
                     TRUE ~ root + general_roots$mean[match(site2, general_roots$site2)])) #+
                     #general_roots$sd[match(site2, general_roots$site2)]))

#write csv for measured and estimated beginning biomass
#write.csv(plant_harvest_est, "begin_estimates_bysp.csv") #saves as csv in current workspace
#write.csv(total_pots, "begin_estimates_bypot.csv") #saves as csv in current workspace

#read in AM data
AMdata <- read_csv("Wigan_myc.csv")

#order variables
plant_harvest$site <- factor(plant_harvest$site, levels = order_sites)
plant_harvest_est$site <- factor(plant_harvest_est$site, levels = order_sites)
plant13C$site <- factor(plant13C$site, levels = order_sites)
summary_plant13C$site <- factor(summary_plant13C$site, levels = order_sites)

AMdata$Site <- factor(AMdata$Site, levels = order_sites)


```

```{r calc plant biodiversity}

#Pivot data to wide format
species_df_wide <- plant_harvest %>%
  dplyr::select(pot, species, shoot) %>% #select only needed columns
  filter(!is.na(species) & !grepl("general_roots", species)) %>% #remove any NA species
  pivot_wider(names_from = species, values_from = shoot) %>% #determines observation name and value columns
  replace(is.na(.), 0) %>% #make any missing values 0
  column_to_rownames(var="pot") #sets observation column as row names

#Calculate species richness, Shannon index, and Simpson index with vegan library
species_richness <- specnumber(species_df_wide) #calculate species richness
shannon <- diversity(species_df_wide, index = "shannon") #calculate Shannon index
simpson <- diversity(species_df_wide, index = "simpson") #calculate Simpson index

#Combine biodiversity metrics into a dataframe
diversity <- data.frame(
  pot = rownames(species_df_wide), #put observations back into column
  species_richness = species_richness,
  shannon = shannon,
  simpson = simpson,
  row.names = NULL
) 

diversity <- diversity %>%
  mutate(site = plant_harvest$site[match(diversity$pot, plant_harvest$pot)]) #add site information back in



```


### Soil organic matter

```{r load data 3.1, include=FALSE}

#read in field soil C/N data
SOM_fielddata <- read_csv("SOM_field.csv") 

#Combine SOM, POM, and MAOM as one observation per plot
SOM <- SOM_fielddata %>%
  filter(Fraction == "SOM") %>%
  rename("SOM_C" = "C_gkg") %>%
  rename("SOM_N" = "N_gkg") %>%
  rename("SOM_CN" = "C_N") %>%
  rename("SOM_C_kgm2_7cm" = "C_kgm2_7cm")

POM <- SOM_fielddata %>%
  filter(Fraction == "POM") %>%
  rename("POM_C" = "C_gkg") %>%
  rename("POM_N" = "N_gkg") %>%
  rename("POM_CN" = "C_N") %>%
  rename("POM_C_kgm2_7cm" = "C_kgm2_7cm")
  
MAOM <- SOM_fielddata %>%
  filter(Fraction == "MAOM") %>%
  rename("MAOM_C" = "C_gkg") %>%
  rename("MAOM_N" = "N_gkg") %>%
  rename("MAOM_CN" = "C_N") %>%
  rename("MAOM_C_kgm2_7cm" = "C_kgm2_7cm")

SOM_fielddata2 <- merge(MAOM,POM, by=c("Site","Management", "Mgmt_num", "Plot_ID"), all.y=TRUE) %>%
  merge(SOM, by=c("Site","Management", "Mgmt_num", "Plot_ID"), all.y=TRUE) %>%
  dplyr::select(Site, Management, Mgmt_num, SOM_C, SOM_N, SOM_CN, POM_C, POM_N, POM_CN, MAOM_C, MAOM_N, MAOM_CN,
                C_gcm3, C_Tha, SOM_C_kgm2_7cm, POM_C_kgm2_7cm, MAOM_C_kgm2_7cm) %>%
  mutate(fMAOM_C = MAOM_C_kgm2_7cm / (POM_C_kgm2_7cm + MAOM_C_kgm2_7cm)) %>% #calculate fMAOM C
  mutate(fMAOM_C_2 = MAOM_C_kgm2_7cm / SOM_C_kgm2_7cm) %>% #calculate fMAOM C
  mutate(fMAOM_N = MAOM_N / (POM_N + MAOM_N)) %>% #calculate fMAOM N
  mutate(fMAOM_N_2 = MAOM_N / SOM_N) #calculate fMAOM N

#summarize SOM data for plots
summary_SOM <- SOM_fielddata2 %>%
  group_by(Site, Management, Mgmt_num) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats

#order variables
SOM_fielddata$Site <- factor(SOM_fielddata$Site, levels = order_sites)
SOM_fielddata$Management <- factor(SOM_fielddata$Management, levels = order_management)
SOM_fielddata2$Site <- factor(SOM_fielddata2$Site, levels = order_sites)
summary_SOM$Site <- factor(summary_SOM$Site, levels = order_sites)
summary_SOM$Management <- factor(summary_SOM$Management, levels = order_management)

```

```{r load data 3.2, include=FALSE}

#read in soil C/N and 13C data
soil_13C_calcs <- read_csv("soil_13C_calcs.csv") %>%
  na.omit() %>%
  mutate(C_gkg = perC_samplemean * 10) %>%
  mutate(N_gkg = perN_samplemean * 10)

#summarize SOM data for plots
summary_soil13C <- soil_13C_calcs %>%
  group_by(Site, Management, Mgmt_num) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats

#make 
TotalC <- soil_13C_calcs %>%
  dplyr::select(Site, C_kgm2_7cm)

#order variables
soil_13C_calcs$Site <- factor(soil_13C_calcs$Site, levels = order_sites)
summary_soil13C$Site <- factor(summary_soil13C$Site, levels = order_sites)

```

```{r load data 3.3 T0, include=FALSE}

#read in soil C/N and 13C data
SOM_13C <- read_csv("SOM_13C.csv")

#Combine SOM, POM, and MAOM as one observation per plot for T0
SOM_T0 <- SOM_13C %>%
  filter(Fraction == "SOM") %>%
  filter(Incubation_T == 0) %>%
  dplyr::select(Site, Management, Mgmt_num, Plot_ID, Incubation_T, per_C, per_N, C_gkg, N_gkg, C_N, d13C, `13C_excess_per`, `15N`, per_POM, per_MAOM, C_kgm2_7cm, pot_C_g, pot_13C_mg) %>% #select columns to keep
  rename("SOM_C_per" = "per_C") %>%
  rename("SOM_N_per" = "per_N") %>%
  rename("SOM_potC_g" = "pot_C_g") %>%
  rename("SOM_C" = "C_gkg") %>%
  rename("SOM_N" = "N_gkg") %>%
  rename("SOM_CN" = "C_N") %>%
  rename("SOM_d13C" = "d13C") %>%
  rename("SOM_13C_excess" = "13C_excess_per") %>%
  rename("SOM_pot13C_mg" = "pot_13C_mg") %>%
  rename("SOM_15N" = "15N") %>%
  rename("SOM_C_kgm2_7cm" = "C_kgm2_7cm")

POM_T0 <- SOM_13C %>%
  filter(Fraction == "POM") %>%
  filter(Incubation_T == 0) %>%
  dplyr::select(Site, Management, Mgmt_num, Plot_ID, Incubation_T, per_C, per_N, C_gkg, N_gkg, C_N, d13C, `13C_excess_per`, `15N`, per_POM, per_MAOM, C_kgm2_7cm, pot_C_g, pot_13C_mg) %>% #select columns to keep
  rename("POM_C_per" = "per_C") %>%
  rename("POM_N_per" = "per_N") %>%
  rename("POM_potC_g" = "pot_C_g") %>%
  rename("POM_C" = "C_gkg") %>%
  rename("POM_N" = "N_gkg") %>%
  rename("POM_CN" = "C_N") %>%
  rename("POM_d13C" = "d13C") %>%
  rename("POM_13C_excess" = "13C_excess_per") %>%
  rename("POM_pot13C_mg" = "pot_13C_mg") %>%
  rename("POM_15N" = "15N") %>%
  rename("POM_C_kgm2_7cm" = "C_kgm2_7cm")
  
MAOM_T0 <- SOM_13C %>%
  filter(Fraction == "MAOM") %>%
  filter(Incubation_T == 0) %>%
  dplyr::select(Site, Management, Mgmt_num, Plot_ID, Incubation_T, per_C, per_N, C_gkg, N_gkg, C_N, d13C, `13C_excess_per`, `15N`, per_POM, per_MAOM, C_kgm2_7cm, pot_C_g, pot_13C_mg) %>% #select columns to keep
  rename("MAOM_C_per" = "per_C") %>%
  rename("MAOM_N_per" = "per_N") %>%
  rename("MAOM_potC_g" = "pot_C_g") %>%
  rename("MAOM_C" = "C_gkg") %>%
  rename("MAOM_N" = "N_gkg") %>%
  rename("MAOM_CN" = "C_N")  %>%
  rename("MAOM_d13C" = "d13C") %>%
  rename("MAOM_13C_excess" = "13C_excess_per") %>%
  rename("MAOM_pot13C_mg" = "pot_13C_mg") %>%
  rename("MAOM_15N" = "15N") %>%
  rename("MAOM_C_kgm2_7cm" = "C_kgm2_7cm")

SOM_13C_T0 <- merge(SOM_T0, POM_T0, by=c("Site","Management", "Mgmt_num", "Plot_ID", "Incubation_T"), 
                    all=TRUE) %>%
  merge(MAOM_T0, by=c("Site","Management", "Mgmt_num", "Plot_ID", "Incubation_T"), all.y=TRUE) %>%
  mutate(fMAOM_C = MAOM_C_kgm2_7cm / (POM_C_kgm2_7cm + MAOM_C_kgm2_7cm)) %>% #calculate fMAOM C
  mutate(fMAOM_C_2 = MAOM_C_kgm2_7cm / SOM_C_kgm2_7cm) %>% #calculate fMAOM C
  mutate(fMAOM_13C = MAOM_pot13C_mg / (MAOM_pot13C_mg + POM_pot13C_mg)) %>% #calculate fMAOM 13C
  mutate(M13C_P13C = MAOM_pot13C_mg / POM_pot13C_mg) %>% #calculate ratio transfer to MAOM vs POM
  mutate(fMAOM_N = MAOM_N / (POM_N + MAOM_N)) %>% #calculate fMAOM N
  mutate(fMAOM_N_2 = MAOM_N / SOM_N) #calculate fMAOM N

SOM_13C_T0 <- SOM_13C_T0 %>%
  mutate(TotalC = soil_13C_calcs$C_kgm2_7cm[match(SOM_13C_T0$Plot_ID, soil_13C_calcs$Plot_ID)]) 

#summarize SOM data for plots for T0
summary_SOM_13C_T0 <- SOM_13C_T0 %>%
  group_by(Site, Management, Mgmt_num) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats

#summarize SOM data for plots and incubation
summary_SOM_13C <- SOM_13C[-137,] %>%
  group_by(Site, Management, Mgmt_num, Incubation_T, Fraction) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats


SOM_13C_T0 <- SOM_13C_T0 %>%
  mutate(AM_hyph_per = AMdata$hyph_percent[match(SOM_13C_T0$Plot_ID, AMdata$Sample)]) %>%
  mutate(AM_hyph_ct = AMdata$Hyphae[match(SOM_13C_T0$Plot_ID, AMdata$Sample)]) %>%
  mutate(AM_arb_per = AMdata$arbuscule_percent[match(SOM_13C_T0$Plot_ID, AMdata$Sample)]) %>%
  mutate(AM_arb_ct = AMdata$Arbuscule[match(SOM_13C_T0$Plot_ID, AMdata$Sample)]) %>%
  mutate(AM_ves_ct = AMdata$Vesicle[match(SOM_13C_T0$Plot_ID, AMdata$Sample)]) %>%
  mutate(AM_absent = AMdata$Nothing[match(SOM_13C_T0$Plot_ID, AMdata$Sample)]) %>%
  mutate(pH = SOM_13C_T0$pH <- site_data$pH[match(SOM_13C_T0$Plot_ID, site_data$Plot_ID)]) %>%
  mutate(pH_range = case_when(pH <= 5 ~ "Very Strongly Acidic",
                                pH > 5 & pH <= 5.5 ~ "Strongly Acidic",
                                pH > 5.5 & pH <= 6 ~ "Moderately Acidic",
                                pH > 6 & pH <= 6.5 ~ "Mildly Acidic",
                                pH > 6.5 & pH <= 7.3 ~ "Neutral",
                                pH > 7.3 ~ "Alkaline"))

SOM_T6 <- SOM_13C %>%
  filter(Incubation_T == 6)

#order variables
SOM_13C$Site <- factor(SOM_13C$Site, levels = order_sites)
SOM_T6$Site <- factor(SOM_T6$Site, levels = order_sites)
SOM_13C_T0$Site <- factor(SOM_13C_T0$Site, levels = order_sites)
summary_SOM_13C$Site <- factor(summary_SOM_13C$Site, levels = order_sites)
summary_SOM_13C_T0$Site <- factor(summary_SOM_13C_T0$Site, levels = order_sites)

```


### 13C respiration and decomposition

```{r load data 4, include=FALSE}

#read in respiration data
resp_min <- read_csv("Wigan_respiration.csv")
resp_full <- read_csv("Wigan_resp_full.csv") %>%
  na.omit() %>% #remove rows with NA values
  mutate(Day = case_when(Date == "1/29/24" ~ 0,
                          Date == "1/30/24" ~ 1,
                          Date == "2/4/24" ~ 6,
                          Date == "2/5/24" ~ 7)) %>%
  mutate(Week = case_when(Day == 0 ~ 0,
                          Day == 1 ~ 0,
                          Day == 6 ~ 1,
                          Day == 7 ~ 1))

#read in incubation data, add beginning, calculate change in biomass
endbiomass <- read_csv("End_incubation_biomass.csv") %>% #load
  mutate(est_beg_drymass = total_pots$root[match(Pot, total_pots$pot)]) %>% #add estimated beginning biomass
  mutate(beg_drymass_perL = est_beg_drymass / 6.1) %>% #calculate dry mass per liter
  mutate(dmass_dT = est_beg_drymass - end_drymass_clean) %>% #calculate change in dry mass
  mutate(dmass_dT_perL = beg_drymass_perL - end_drymass_perL_clean) %>% #calculate change in dry mass per liter
  mutate(decomp_per_unadj = dmass_dT / est_beg_drymass) %>% #calculate decomposition percentage - not adjusting volume
  mutate(decomp_per_adj = dmass_dT_perL / beg_drymass_perL) %>% #calculate decomposition percentage - adjusting volume
  mutate(years =
           case_when(Management == "Unmanaged" ~ 0,
                     Management == "4 Years" ~ 4,
                     Management == "15 Years" ~ 15,
                     Management == "23 Years" ~ 23))

#Add data
endbiomass <- endbiomass %>%
  mutate(AM_hyph_per = AMdata$hyph_percent[match(endbiomass$Pot, AMdata$Sample)]) %>%
  mutate(AM_hyph_ct = AMdata$Hyphae[match(endbiomass$Pot, AMdata$Sample)]) %>%
  mutate(AM_arb_per = AMdata$arbuscule_percent[match(endbiomass$Pot, AMdata$Sample)]) %>%
  mutate(AM_arb_ct = AMdata$Arbuscule[match(endbiomass$Pot, AMdata$Sample)]) %>%
  mutate(AM_ves_ct = AMdata$Vesicle[match(endbiomass$Pot, AMdata$Sample)]) %>%
  mutate(AM_absent = AMdata$Nothing[match(endbiomass$Pot, AMdata$Sample)]) %>%
  mutate(pH = SOM_13C_T0$pH <- site_data$pH[match(endbiomass$Pot, site_data$Plot_ID)]) %>%
  mutate(pH_range = case_when(pH <= 5 ~ "Very Strongly Acidic",
                                pH > 5 & pH <= 5.5 ~ "Strongly Acidic",
                                pH > 5.5 & pH <= 6 ~ "Moderately Acidic",
                                pH > 6 & pH <= 6.5 ~ "Mildly Acidic",
                                pH > 6.5 & pH <= 7.3 ~ "Neutral",
                                pH > 7.3 ~ "Alkaline"))


#summarise
summary_endbio <- endbiomass %>%
  dplyr::select(Site, Management, years, rocks_debris, end_drymass_clean, end_drymass_perL_clean, est_beg_drymass, beg_drymass_perL, decomp_per_adj) %>%
  group_by(Site, Management, years) %>%
  summarise(across(where(is.numeric), summary_stats)) #run all summary stats

#order variables for each
resp_min$Site <- factor(resp_min$Site, levels = order_sites)
resp_full$Site <- factor(resp_full$Site, levels = order_sites)
endbiomass$Site <- factor(endbiomass$Site, levels = order_sites)
summary_endbio$Site <- factor(summary_endbio$Site, levels = order_sites)

```

### UK soil data to compare

Manning, Peter; De Vries, Franciska T.; Tallowin, Jerry R. B. et al. (2016). Data from: Simple measures of climate, soil properties and plant traits predict national scale grassland soil carbon stocks [Dataset]. Dryad. https://doi.org/10.5061/dryad.s7867

```{r load UK data}

#load dataset UK grasslands - from DRYAD repository, Manning et al. 2016
UK_soil <- read_csv("Manning2015_dryad.csv") %>%
  filter(type == "U") %>% #use only unimproved sites
  mutate(POM_C = C250 + C50_250) %>% #match to Wigan data
  mutate(MAOM_C = C50_0.45) %>% #match to Wigan data
  rename("pH" = "PH") %>% #match case with Wigan data
  mutate(vegetation = str_replace(vegetation, "wet grassland", "Wet Grassland")) %>% #match naming with Wigan data
  mutate(vegetation = str_replace(vegetation, "Acid grass", "Acid Grass")) %>% #match naming with Wigan data
  dplyr::select(vegetation, type, site, pH, TotalC, POM_C, MAOM_C) %>%
  mutate_at(vars(pH, TotalC, POM_C, MAOM_C), as.numeric) %>%
  na.omit() #only include complete


summary_UKsoils <- UK_soil %>%
  group_by(vegetation) %>%
  summarise(across(pH:MAOM_C, list(mean = ~ mean(., na.rm=TRUE),
                                 sd = ~ sd(., na.rm=TRUE),
                                 se = ~ std.error(., na.rm=TRUE))))


soilC_compare <- SOM_13C_T0 %>%
  dplyr::select(Site, TotalC, POM_C_kgm2_7cm, MAOM_C_kgm2_7cm, pH) %>%
  rename("site" = "Site") %>%
  rename("POM_C" = "POM_C_kgm2_7cm") %>%
  rename("MAOM_C" = "MAOM_C_kgm2_7cm") %>%
  mutate(type = "R") %>% #R for restored
  mutate(vegetation = case_when(grepl("Bickershaw 1", site) #use grepl to find a substring within a column
															  ~ "Bickershaw Wet Grassland Uncut 4 Yrs", #set value in new column 
															  grepl("Bickershaw 2", site) ~ "Bickershaw Wet Grassland 4 Yrs", 
															  grepl("Amberswood", site) ~ "Amberswood Mesotrophic 15 Yrs", 
															  grepl("Flashes", site) ~ "Flashes Mesotrophic 23 Yrs")) %>% 
  bind_rows(UK_soil) %>%
  mutate(pH_range = case_when(pH <= 5 ~ "Very Strongly Acidic",
                                pH > 5 & pH <= 5.5 ~ "Strongly Acidic",
                                pH > 5.5 & pH <= 6 ~ "Moderately Acidic",
                                pH > 6 & pH <= 6.5 ~ "Mildly Acidic",
                                pH > 6.5 & pH <= 7.3 ~ "Neutral",
                                pH > 7.3 ~ "Alkaline")) #%>%
  #mutate(name = fct_relevel(vegetation, #change ordering manually
  #                          "Acid grass", "Calcicolous", "Mesotrophic", "wet grassland", "restored"))


#order variables
soilC_compare$vegetation <- factor(soilC_compare$vegetation, 
                                   levels = c("Acid Grass", "Calcicolous", "Mesotrophic", "Wet Grassland", 
                                              "Bickershaw Wet Grassland Uncut 4 Yrs",
                                              "Bickershaw Wet Grassland 4 Yrs",
                                              "Amberswood Mesotrophic 15 Yrs",
                                              "Flashes Mesotrophic 23 Yrs"))
soilC_compare$pH_range <- factor(soilC_compare$pH_range, 
                                   levels = c("Very Strongly Acidic", "Strongly Acidic", "Moderately Acidic",
                                              "Mildly Acidic", "Neutral", "Alkaline"))
soilC_compare$type <- factor(soilC_compare$type, levels = c("U", "R")) 

```

# Stats and Charts

## Soil properties
```{r soil char}

#soil pH differences
shapiro.test(resid(lm(pH ~ Management, data = site_data))) #check normality of residuals - not normal
kruskal.test(pH ~ Management, data = site_data) #Kruskal-Wallis test - significant
dunn.test(site_data$pH, site_data$Management, method = "bonferroni", kw=TRUE) #23 yr vs. 4 and 0, 15 vs. 4

#soil bulk density differences
shapiro.test(resid(lm(BD ~ Management, data = site_data))) #check normality of residuals - normal
summary(aov(BD ~ Management, data = site_data)) #ANOVA - not significant


```

## Soil metals

```{r metals comparisons}

#pairwise comparisons
#function 1 evaluate_group_sizes
#function2 differences_function

# Define what to analyze
factors <- "Site" #predictor variable
continuous_vars <- c("ppm_Al", "ppm_B", "ppm_Ba", "ppm_Ca1", "ppm_Ca2", "ppm_Cd", "ppm_Co", "ppm_Cr", "ppm_Cu",
                     "ppm_Fe", "ppm_K", "ppm_Li", "ppm_Mg", "ppm_Mn", "ppm_Na", "ppm_Ni", "ppm_P1", "ppm_P2",
                     "ppm_Pb", "ppm_S1", "ppm_S2", "ppm_Si", "ppm_Sr1", "ppm_Sr2", "ppm_Zn") #response variables
data <- site_data

# Evaluate group sizes
evaluate_group_sizes(data, factors, continuous_vars) 

# Run the analysis
metals_diff <- differences_function(data, factors, continuous_vars)


```

```{r metals chart}

soil_metals$Site <- factor(soil_metals$Site, levels = order_sites)
soil_metals$Management <- factor(soil_metals$Management, levels = order_management)

#plot metals
ggplot(data = soil_metals, aes(x=nutrient, y=ppm, fill=nutrient)) + 
  geom_boxplot() +
  facet_wrap(~Site, scales = "free_x") + #create a separate plot for each site
  scale_fill_manual(values=colors_ramp) + #define colors to use
  ggtitle("Soil Metals per site") + #add a graph title
  xlab("Site") + ylab("ppm") + #define/change X and Y labels
  theme(axis.text.x = element_text(angle=90), #set angle of variable labels
        panel.grid.major.y = element_line(color = "grey40", size = 0.25)) + 
  my_theme

#save figure
ggsave('metals_all.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications


```

```{r metals chart 2}

#plot metals
soil_metals %>%
  filter(nutrient == "Ca1" | nutrient == "Li" | nutrient == "Mg" | 
           nutrient == "Pb" | nutrient == "S1") %>% #filter for significant nutrients
ggplot(aes(x=nutrient, y=ppm, fill=nutrient)) + 
  geom_boxplot() +
  facet_wrap(~Site, scales = "free_x") + #create a separate plot for each site
  scale_fill_manual(values=colors5) + #define colors to use
  #ggtitle("Soil Metals per site") + #add a graph title
  xlab("Site") + ylab("Nutrient (ppm)") + #define/change X and Y labels
  theme(axis.text.x = element_text(angle=90), #set angle of variable labels
        panel.grid.major.y = element_line(color = "grey40", size = 0.25)) + 
  my_theme

#save figure
ggsave('metals_sig.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```


```{r metals chart 3}

#plot metals
soil_metals %>%
  filter(nutrient == "Ca1" | nutrient == "Li" | nutrient == "Mg" | 
           nutrient == "Pb" | nutrient == "S1") %>% #filter for significant nutrients
ggplot(aes(x=Site, y=ppm, fill=Site)) + 
  geom_boxplot() +
  facet_wrap(~nutrient, scales = "free", nrow=1) + #create a separate plot for each site
  scale_fill_manual(values=colors5) + #define colors to use
  #ggtitle("Soil Metals per site") + #add a graph title
  xlab("Site") + ylab("Nutrient (ppm)") + #define/change X and Y labels
  theme(axis.text.x = element_text(angle=90), #set angle of variable labels
        panel.grid.major.y = element_line(color = "grey40", size = 0.25)) + 
  scale_x_discrete(labels = "", "", "", "") + #remove x axis labels
  my_theme +
  theme(legend.position='bottom') #remove legend

#save figure
ggsave('metals_sig2.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 8, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```


## FTIR compounds

```{r FTIR stats}

#define what to analyze for differences
factors <- "site" #predictor variable
continuous_vars <- c("x907", "x1002", "x1402", "x1630", "x794", "x2846", "x2937", "x3620") #response variables
data <- FTIR %>%
  clean_names()

#evaluate group sizes
evaluate_group_sizes(data, factors, continuous_vars) 

#eun the analysis
FTIR_diff <- differences_function(data, factors, continuous_vars)

#aromatic:aliphatic ratio
FTIR_stbl_lbl <- FTIR %>%
  clean_names() %>%
  group_by(site) %>%
  mutate(ar_al = (x794)/(x1402)) %>%
  mutate(stbl_lbl = (x794+x2846+x2937)/(x907+x1002+x1402)) 

FTIR_summary <- FTIR_stbl_lbl %>%
  summarise(across(ar_al:stbl_lbl, list(mean = ~ mean(., na.rm=TRUE),
                                 sd = ~ sd(., na.rm=TRUE),
                                 se = ~ std.error(., na.rm=TRUE))))

shapiro.test(resid(lm(ar_al ~ site, data = FTIR_stbl_lbl))) #check normality of residuals - normal
summary(aov(ar_al ~ site, data = FTIR_stbl_lbl)) #ANOVA - significant
TukeyHSD(aov(ar_al ~ site, data = FTIR_stbl_lbl)) #Flashes vs. all, Amberswood vs. Bickershaw 1

```

```{r FTIR plot}

#plot order
FTIR_long$wavelength <- factor(FTIR_long$wavelength, 
                             levels = c("907", "1002", "1402", "1630", "794", "2846", "2937", "3620"))
#facet_wrap plot
FTIR_long %>%
ggplot(aes(x=Site, y=absorbance, fill=Site)) + 
  geom_boxplot() +
  facet_wrap(~wavelength, scales = "free", nrow=2) + #create a separate plot for each site
  scale_fill_manual(values=colors9) + #define colors to use
  #ggtitle("Soil carbon compounds per site") + #add a graph title
  xlab("Site") + ylab(expression("Absorbance (cm"^-1*")")) + #define/change X and Y labels
  theme(#axis.text.x = element_text(angle=90), #set angle of variable labels
        panel.grid.major.y = element_line(color = "grey40", size = 0.25)) + 
  scale_x_discrete(labels = "", "", "", "") + #remove x axis labels
  my_theme +
  theme(legend.position='bottom') #remove legend

#save figure
ggsave('FTIR_boxplot.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 2000, height = 1200, units = "px", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications


```

```{r FTIR PCA}

#PCA
FTIR_pca <- prcomp(FTIR_standardized_num) #perform PCA on standardized variables

summary(FTIR_pca) #view importance of principal components and proportion of variances

FTIR_pca$rotation #view loadings from the standardized data
screeplot(FTIR_pca) # diagnostic plot to make sure principal components are in order

#basic plot
biplot(FTIR_pca)

#customized plot
autoplot(FTIR_pca, data = FTIR_standardized, colour = 'Site', fill = 'Site',
         loadings = TRUE, loadings.colour = 'black',
         loadings.label = TRUE, loadings.label.size = 3,
         frame = TRUE, frame.type = 'norm', frame.colour = 'Site', frame.fill = 'Site',
         shape = 21) +
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  my_theme +
  ggtitle("Principal Components of FTIR Wavelengths Across Sites") 

#save figure
ggsave('FTIR_PCA.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder within working directory
       width = 6, height = 4, units = "in", #set dimensions for figure
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications


```

## Soil C and N

### Sam Kimpton's data

```{r SOC chart - Kimpton}

#SOM C bar chart with error bars
ggplot(data = summary_SOM, aes(x=Site, y=SOM_C_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=SOM_C_mean-SOM_C_std.error, ymax=SOM_C_mean+SOM_C_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Soil Organic Matter C per site") + #add a graph title
  xlab("Site") + ylab("C (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOC_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r SOC ha - Kimpton}

#SOM C bar chart with error bars per hectare
ggplot(data = summary_SOM, aes(x=Site, y=C_Tha_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=(C_Tha_mean-C_Tha_std.error), ymax=(C_Tha_mean+C_Tha_std.error)),
  color = "grey10", width = 0.7, position = "dodge") + 
  scale_y_continuous(breaks = pretty(c(0,300), n = 6)) + #set limits and change number of tick marks on y axis
  ggtitle("Soil Organic Matter C per hectare") + #add a graph title
  xlab("Site") + ylab("C in top 10cm (T/ha)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOC_ha_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r SOM-N - Kimpton}

#SOM N bar chart with error bars
ggplot(data = summary_SOM, aes(x=Site, y=SOM_N_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=SOM_N_mean-SOM_N_std.error, ymax=SOM_N_mean+SOM_N_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Soil Organic Matter N per site") + #add a graph title
  xlab("Site") + ylab("N (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOM_N_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r SOM-CN - Kimpton}

#SOM CN boxplot
ggplot(data = SOM_fielddata2) + 
	geom_boxplot(aes(x=Site, y=SOM_CN, fill=Site)) +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("Soil Organic Matter C:N per site") + #add a graph title
  xlab("Site") + ylab("C:N") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOM_CN_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r combined C plots - Kimpton}

#combined boxplots with error bars
ggplot(data = SOM_fielddata) + 
	geom_boxplot(aes(x=Site, y=C_gkg, fill=Fraction)) +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("Soil Organic Matter C per site") + #add a graph title
  xlab("Site") + ylab("C (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOMC_all_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r combined N plots - Kimpton}

#combined boxplots with error bars
ggplot(data = SOM_fielddata) + 
	geom_boxplot(aes(x=Site, y=N_gkg, fill=Fraction)) +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("Soil Organic Matter N per site") + #add a graph title
  xlab("Site") + ylab("N (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOM_N_all_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r POM-C - Kimpton}

#POM C bar chart with error bars
ggplot(data = summary_SOM, aes(x=Site, y=POM_C_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=POM_C_mean-POM_C_std.error, ymax=POM_C_mean+POM_C_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Particulate Organic Matter C per site") + #add a graph title
  xlab("Site") + ylab("C (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('POM_Cbar_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r POM-N - Kimpton}

#POM N bar chart with error bars
ggplot(data = summary_SOM, aes(x=Site, y=POM_N_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=POM_N_mean-POM_N_std.error, ymax=POM_N_mean+POM_N_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Particulate Organic Matter N per site") + #add a graph title
  xlab("Site") + ylab("N (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('POM_Nbar_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r MAOM-C - Kimpton}

#MAOM C bar chart with error bars
ggplot(data = summary_SOM, aes(x=Site, y=MAOM_C_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=MAOM_C_mean-MAOM_C_std.error, ymax=MAOM_C_mean+MAOM_C_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Mineral Associated Organic Matter C per site") + #add a graph title
  xlab("Site") + ylab("C (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('MAOM_Cbar_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r MAOM-N - Kimpton}

#MAOM N bar chart with error bars
ggplot(data = summary_SOM, aes(x=Site, y=MAOM_N_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=MAOM_N_mean-MAOM_N_std.error, ymax=MAOM_N_mean+MAOM_N_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Mineral Associated Organic Matter N per site") + #add a graph title
  xlab("Site") + ylab("N (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('MAOM_Nbar_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r fMAOM-C - Kimpton}

#fMAOM C box chart
ggplot(data = SOM_fielddata2, aes(x=Site, y=fMAOM_C, fill=Site)) + 
	geom_boxplot() +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("fMAOM C per site") + #add a graph title
  xlab("Site") + ylab("Portion of SOC stored in MAOM") + #define/change X and Y labels
  scale_y_continuous(labels = scales::percent) + #change y axis to percent
  my_theme

#save figure
ggsave('fMAOM_C_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r fMAOM-N - Kimpton}

#fMAOM N box chart
ggplot(data = SOM_fielddata2, aes(x=Site, y=fMAOM_N, fill=Site)) + 
	geom_boxplot() +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("fMAOM N per site") + #add a graph title
  xlab("Site") + ylab("Portion of SOC stored in MAOM") + #define/change X and Y labels
  scale_y_continuous(labels = scales::percent) + #change y axis to percent
  my_theme

#save figure
ggsave('fMAOM_N_Kimpton.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

### Updated data

```{r SOC chart - sample means}

#SOM C bar chart with error bars
summary_soil13C %>%
ggplot(aes(x=Site, y=C_gkg_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=C_gkg_mean-C_gkg_std.error, ymax=C_gkg_mean+C_gkg_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Soil Organic Matter C per site") + #add a graph title
  xlab("") + ylab("C (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOC_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r SOC chart 2 - sample means}

#SOM C bar chart with error bars
summary_soil13C %>%
ggplot(aes(x=Site, y=C_kgm2_7cm_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=C_kgm2_7cm_mean-C_kgm2_7cm_std.error, 
                    ymax=C_kgm2_7cm_mean+C_kgm2_7cm_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Soil Organic Matter C per site") + #add a graph title
  xlab("") + ylab("SOC (kg/m2 in top 7cm)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOC_kgm2_means.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r SOM-N - sample means}

#SOM N bar chart with error bars
summary_soil13C %>%
ggplot(aes(x=Site, y=N_gkg_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=N_gkg_mean-N_gkg_std.error, ymax=N_gkg_mean+N_gkg_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Soil Organic Matter N per site") + #add a graph title
  xlab("") + ylab("N (g/kg)") + #define/change X and Y labels
  scale_y_continuous(breaks = pretty) + #set limits and change number of tick marks on y axis
  my_theme

#save figure
ggsave('SOM_N_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r SOM-CN - Barker}

#SOM CN boxplot
ggplot(data = soil_13C_calcs) + 
	geom_boxplot(aes(x=Site, y=C_N, fill=Site)) +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("Soil Organic Matter C:N per site") + #add a graph title
  xlab("Site") + ylab("C:N") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOM_CN_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r combined C plots - Barker}

#combined boxplots with error bars
ggplot(data = SOM_13C) + 
	geom_boxplot(aes(x=Site, y=C_gkg, fill=Fraction)) +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("Soil Organic Matter C per site") + #add a graph title
  xlab("Site") + ylab("C (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOMC_all_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r combined N plots - Barker}

#combined boxplots with error bars
ggplot(data = SOM_13C) + 
	geom_boxplot(aes(x=Site, y=N_gkg, fill=Fraction)) +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("Soil Organic Matter N per site") + #add a graph title
  xlab("Site") + ylab("N (g/kg)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('SOM_N_all_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r POM-C - Barker}

#POM C bar chart with error bars
ggplot(data = summary_SOM_13C_T0, aes(x=Site, y=POM_C_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=POM_C_mean-POM_C_std.error, ymax=POM_C_mean+POM_C_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Particulate Organic Matter C per site") + #add a graph title
  xlab("") + ylab("POM-C (g C/kg soil)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('POM_Cbar_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r POM-N - Barker}

#POM N bar chart with error bars
ggplot(data = summary_SOM_13C_T0, aes(x=Site, y=POM_N_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=POM_N_mean-POM_N_std.error, ymax=POM_N_mean+POM_N_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Particulate Organic Matter N per site") + #add a graph title
  xlab("Site") + ylab("POM-N (g N/kg soil)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('POM_Nbar_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r MAOM-C - Barker}

#MAOM C bar chart with error bars
ggplot(data = summary_SOM_13C_T0, aes(x=Site, y=MAOM_C_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=MAOM_C_mean-MAOM_C_std.error, ymax=MAOM_C_mean+MAOM_C_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Mineral Associated Organic Matter C per site") + #add a graph title
  xlab("Site") + ylab("MAOM-C (g C/kg soil)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('MAOM_Cbar_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r MAOM-N - Barker}

#MAOM N bar chart with error bars
ggplot(data = summary_SOM_13C_T0, aes(x=Site, y=MAOM_N_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=MAOM_N_mean-MAOM_N_std.error, ymax=MAOM_N_mean+MAOM_N_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Mineral Associated Organic Matter N per site") + #add a graph title
  xlab("Site") + ylab("MAOM-N (g N/kg soil)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('MAOM_Nbar_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r fMAOM-C - Barker}

#fMAOM C box chart
ggplot(data = SOM_13C_T0, aes(x=Site, y=fMAOM_C, fill=Site)) + 
	geom_boxplot() +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("fMAOM C per site") + #add a graph title
  xlab("Site") + ylab("Portion of SOC stored in MAOM") + #define/change X and Y labels
  scale_y_continuous(labels = scales::percent) + #change y axis to percent
  my_theme

#save figure
ggsave('fMAOM_C_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r fMAOM-N - Barker}

#fMAOM N box chart
ggplot(data = SOM_13C_T0, aes(x=Site, y=fMAOM_N, fill=Site)) + 
	geom_boxplot() +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("fMAOM N per site") + #add a graph title
  xlab("Site") + ylab("Portion of SON stored in MAOM") + #define/change X and Y labels
  scale_y_continuous(labels = scales::percent) + #change y axis to percent
  my_theme

#save figure
ggsave('fMAOM_N_KB.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

### UK unimproved grassland soils C
```{r SOC UK}

#SOM C bar chart with error bars
ggplot(data = summary_UKsoils, aes(x=vegetation, y=TotalC_mean, fill=vegetation)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors9) + #define colors to use
  geom_errorbar(aes(ymin=TotalC_mean-TotalC_se, 
                    ymax=TotalC_mean+TotalC_se),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Soil Organic Matter C across UK grasslands") + #add a graph title
  xlab("Grassland Type") + ylab("SOC (kg/m2 in top 7cm)") + #define/change X and Y labels
  my_theme


#save figure
ggsave('SOC_Manning.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications


```

```{r compare soil C 1}

ggplot(data = soilC_compare) + 
	geom_boxplot(aes(x=vegetation, y=TotalC, fill=vegetation)) +
  scale_fill_manual(values=colors9) + #define colors to use
  ggtitle("Soil Organic Matter C across the UK") + #add a graph title
  xlab("") + ylab(expression("SOC (kg/m"^2*" in top 7cm)")) + #define/change X and Y labels
  scale_x_discrete(labels = c("", "", "", "", "", "", "", "")) + #remove vegetation labels from x axis
  my_theme 

#save figure
ggsave('SOC_comparison.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 2000, height = 1000, units = "px", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r compare soil C 2}

ggplot(data = soilC_compare) + 
	geom_boxplot(aes(x=type, y=TotalC, fill=vegetation)) +
  scale_fill_manual(values=colors9) + #define colors to use
  ggtitle("Soil Organic Matter C across the UK") + #add a graph title
  xlab("") + ylab(expression("SOC (kg/m"^2*" in top 7cm)")) + #define/change X and Y labels
  scale_x_discrete(labels = c("Unimproved Grasslands", "Restored Grasslands")) + #remove vegetation labels from x axis
  my_theme 

#save figure
ggsave('SOC_comparison2.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 2000, height = 1000, units = "px", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```


```{r compare POM C}

ggplot(data = soilC_compare) + 
	geom_boxplot(aes(x=type, y=POM_C, fill=vegetation)) +
  scale_fill_manual(values=colors9) + #define colors to use
  #ggtitle("Particulate Organic Matter C across the UK") + #add a graph title
  xlab("") + ylab(expression("POM-C (kg/m"^2*" in top 7cm)")) + #define/change X and Y labels
  scale_x_discrete(labels = c("Unimproved Grasslands", "Restored Grasslands")) + #remove vegetation labels from x axis
  my_theme 

#save figure
ggsave('POM-C_comparison.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 2000, height = 800, units = "px", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r compare MAOM C}

ggplot(data = soilC_compare) + 
	geom_boxplot(aes(x=type, y=MAOM_C, fill=vegetation)) +
  scale_fill_manual(values=colors9) + #define colors to use
  #ggtitle("Mineral Associated Organic Matter C across the UK") + #add a graph title
  xlab("") + ylab(expression("MAOM-C (kg/m"^2*" in top 7cm)")) + #define/change X and Y labels
  scale_x_discrete(labels = c("Unimproved Grasslands", "Restored Grasslands")) + #remove vegetation labels from x axis
  my_theme 

#save figure
ggsave('MAOM-C_comparison.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 2000, height = 800, units = "px", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

## SOM Stats

### Compare with UK-wide unimproved grasslands soil C
```{r SOC compare}

shapiro.test(resid(lm(TotalC ~ type, data = soilC_compare))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(TotalC) ~ type, data = soilC_compare))) #not normal, closer
summary(aov(log(TotalC) ~ type, data = soilC_compare)) #very sig

shapiro.test(resid(lm(TotalC ~ vegetation, data = soilC_compare))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(TotalC) ~ vegetation, data = soilC_compare))) #not normal
summary(aov(log(TotalC) ~ vegetation, data = soilC_compare)) #very sig
TukeyHSD(aov(log(TotalC) ~ vegetation, data = soilC_compare)) #all restored > acid, Amber & Flashes > Calc, all rest > Meso, all rest > wet

shapiro.test(resid(lm(POM_C ~ type, data = soilC_compare))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(POM_C) ~ type, data = soilC_compare))) #not normal, closer
summary(aov(log(POM_C) ~ type, data = soilC_compare)) #sig

shapiro.test(resid(lm(POM_C ~ vegetation, data = soilC_compare))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(POM_C) ~ vegetation, data = soilC_compare))) #normal
summary(aov(log(POM_C) ~ vegetation, data = soilC_compare)) #very sig
TukeyHSD(aov(log(POM_C) ~ vegetation, data = soilC_compare)) #Flashes > all, Bick 1 < calc, Amber > Meso

shapiro.test(resid(lm(MAOM_C ~ type, data = soilC_compare))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(MAOM_C) ~ type, data = soilC_compare))) #not normal, closer
summary(aov(log(MAOM_C) ~ type, data = soilC_compare)) #sig

shapiro.test(resid(lm(MAOM_C ~ vegetation, data = soilC_compare))) #check normality of residuals - not normal but close
shapiro.test(resid(lm(log(MAOM_C) ~ vegetation, data = soilC_compare))) #worse - might use another transform or keep raw
summary(aov(MAOM_C ~ vegetation, data = soilC_compare)) #very sig
TukeyHSD(aov(MAOM_C ~ vegetation, data = soilC_compare)) #all rest > all non
  
```

### Compare sites at Wigan
```{r SOM stats Wigan}

#Note this is redundant from above across all UK, but just uses Wigan dataset. Keep for now.

shapiro.test(resid(lm(SOM_C ~ Site, data = SOM_13C_T0))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(SOM_C) ~ Site, data = SOM_13C_T0))) 
summary(aov(log(SOM_C) ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(log(SOM_C) ~ Site, data = SOM_13C_T0)) #Flashes vs. all, Bick 1 vs. Amber

shapiro.test(resid(lm(SOM_N ~ Site, data = SOM_13C_T0))) #normal
summary(aov(SOM_N ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(SOM_N ~ Site, data = SOM_13C_T0)) #Flashes vs. all

shapiro.test(resid(lm(POM_C ~ Site, data = SOM_13C_T0))) #normal
summary(aov(POM_C ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(POM_C ~ Site, data = SOM_13C_T0)) #Flashes vs. all

shapiro.test(resid(lm(POM_N ~ Site, data = SOM_13C_T0))) #not normal
shapiro.test(resid(lm(log(POM_N) ~ Site, data = SOM_13C_T0))) #normal
summary(aov(log(POM_N) ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(log(POM_N) ~ Site, data = SOM_13C_T0)) #Flashes vs. all, Bick 1 vs. all

shapiro.test(resid(lm(MAOM_C ~ Site, data = SOM_13C_T0))) #normal
summary(aov(MAOM_C ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(MAOM_C ~ Site, data = SOM_13C_T0)) #Flashes vs. all

shapiro.test(resid(lm(MAOM_N ~ Site, data = SOM_13C_T0))) #normal
summary(aov(MAOM_N ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(MAOM_N ~ Site, data = SOM_13C_T0)) #Flashes vs. Bick 1, Amber

shapiro.test(resid(lm(fMAOM_C ~ Site, data = SOM_13C_T0))) #normal
summary(aov(fMAOM_C ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(fMAOM_C ~ Site, data = SOM_13C_T0)) #Flashes vs. Bick 1 & 2, Amber vs. Bick 1

shapiro.test(resid(lm(fMAOM_N ~ Site, data = SOM_13C_T0))) #normal
summary(aov(fMAOM_N ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(fMAOM_N ~ Site, data = SOM_13C_T0)) #Flashes vs. Bick 1 & 2, Amber vs. Bick 1

```

```{r SOM vs pH}

shapiro.test(resid(lm(SOM_C ~ Site * site_data$pH, data = SOM_13C_T0))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(SOM_C) ~ Site * site_data$pH, data = SOM_13C_T0))) #check normality of residuals - normal
summary(aov(log(SOM_C) ~ Site * site_data$pH, data = SOM_13C_T0)) #Site only

shapiro.test(resid(lm(SOM_C_kgm2_7cm ~ Site * site_data$pH, data = SOM_13C_T0))) #not normal
shapiro.test(resid(lm(log(SOM_C_kgm2_7cm) ~ Site * site_data$pH, data = SOM_13C_T0))) #normal
summary(aov(log(SOM_C_kgm2_7cm) ~ Site * site_data$pH, data = SOM_13C_T0)) #Site only

shapiro.test(resid(lm(SOM_N ~ Site * site_data$pH, data = SOM_13C_T0))) #check normality of residuals - normal
summary(aov(SOM_N ~ Site * site_data$pH, data = SOM_13C_T0)) #Site only

```



### Linear Mixed Models explaining soil C

#### By soil pH

```{r SOC compare pH}

shapiro.test(resid(lm(SOM_C ~ pH_range, data = SOM_13C_T0))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(SOM_C) ~ pH_range, data = SOM_13C_T0))) #not normal but closer
summary(aov(log(SOM_C) ~ pH_range, data = SOM_13C_T0)) #sig
summary(aov(log(SOM_C) ~ pH, data = SOM_13C_T0)) #sig

Anova(lmer(log(SOM_C) ~ pH_range +
             (1|Site), #site random
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit), type=3) #not sig

shapiro.test(resid(lm(POM_C ~ pH_range, data = SOM_13C_T0))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(POM_C) ~ pH_range, data = SOM_13C_T0))) #normal
summary(aov(log(POM_C) ~ pH_range, data = SOM_13C_T0)) #sig
summary(aov(log(POM_C) ~ pH, data = SOM_13C_T0)) #sig

Anova(lmer(log(POM_C) ~ pH_range + (1|Site), #site random
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit), type=3) #pH barely sig

POM_C_pH_lme <- lmer(POM_C ~ pH_range + (1|Site), 
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit)
summary(POM_C_pH_lme) #random accounts for most variance
shapiro.test(resid(POM_C_pH_lme)) #normal
performance::r2(POM_C_pH_lme, multivariate=TRUE) #calculate R-squared for model fit
MuMIn::AICc(POM_C_pH_lme, k=2) #calculate AICc for for model comparison with small sample sizes
TukeyHSD(aov(POM_C ~ pH_range + Site, data = SOM_13C_T0)) #acidic lower

shapiro.test(resid(lm(MAOM_C ~ pH_range, data = SOM_13C_T0))) #check normality of residuals - normal
summary(aov(MAOM_C ~ pH_range, data = SOM_13C_T0)) #very sig
summary(aov(MAOM_C ~ pH, data = SOM_13C_T0)) #very sig

Anova(lmer(log(MAOM_C) ~ pH_range + (1|Site), #site random
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit), type=3) #sig

MAOM_C_pH_lme <- lmer(MAOM_C ~ pH_range + (1|Site), 
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit)
summary(MAOM_C_pH_lme) #random accounts for less
shapiro.test(resid(MAOM_C_pH_lme)) #normal
performance::r2(MAOM_C_pH_lme, multivariate=TRUE) #calculate R-squared for model fit
MuMIn::AICc(MAOM_C_pH_lme, k=2) #calculate AICc for for model comparison with small sample sizes
TukeyHSD(aov(MAOM_C ~ pH_range + Site, data = SOM_13C_T0)) #acidic lower
  
```

```{r SOM C vs pH plot}

#SOM-C plot
ggplot(data = SOM_13C_T0, aes(x=pH, y=SOM_C)) + 
  geom_point(aes(color = factor(Site))) + #add points to plot
  geom_smooth(method="lm", se=TRUE, color=plotline_color, fill=fill_color) +
  scale_color_manual(values = colors) + #define colors to use
  my_theme +
  xlab("pH") + ylab("SOM-C (g C/kg soil)") #define/change X and Y labels

ggsave('SOM_C_bypH.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

```

```{r POM C vs pH plot}

#POM-C plot
ggplot(data = SOM_13C_T0, aes(x=pH, y=POM_C)) + 
  geom_point(aes(color = factor(Site))) + #add points to plot
  geom_smooth(method="lm", se=TRUE, color=plotline_color, fill=fill_color) +
  scale_color_manual(values = colors) + #define colors to use
  my_theme +
  xlab("pH") + ylab("POM-C (g C/kg soil)") #define/change X and Y labels

ggsave('POM_C_bypH.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

```

```{r MAOM C vs pH plot}

#SOM-C plot
ggplot(data = SOM_13C_T0, aes(x=pH, y=MAOM_C)) + 
  geom_point(aes(color = factor(Site))) + #add points to plot
  geom_smooth(method="lm", se=TRUE, color=plotline_color, fill=fill_color) +
  scale_color_manual(values = colors) + #define colors to use
  my_theme +
  xlab("pH") + ylab("MAOM-C (g C/kg soil)") #define/change X and Y labels

ggsave('MAOM_C_bypH.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

```


#### By mycorrhizal association
```{r SOM vs AM}

shapiro.test(resid(lm(SOM_C ~ AM_hyph_per + AM_arb_per, data = SOM_13C_T0))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(SOM_C) ~ AM_hyph_per + AM_arb_per, data = SOM_13C_T0))) #check normality of residuals - normal
summary(aov(log(SOM_C) ~ AM_hyph_per + AM_arb_per, data = SOM_13C_T0)) #hyphae sig

Anova(lmer(log(SOM_C) ~ AM_hyph_ct + AM_arb_ct + AM_ves_ct + AM_absent +
             (1|Site), #site random
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit), type=3) #none

Anova(lmer(SOM_C ~ AM_hyph_per + AM_arb_per + AM_ves_ct +
             (1|Site), #site random
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit), type=3) #none

shapiro.test(resid(lm(SOM_N ~ AM_hyph_per + AM_arb_per + AM_ves_ct, data = SOM_13C_T0))) #check normality of residuals - normal
summary(aov(SOM_N ~ AM_hyph_per + AM_arb_per + AM_ves_ct, data = SOM_13C_T0)) #hyphae

Anova(lmer(SOM_N ~ AM_hyph_per + AM_arb_per + AM_ves_ct + (1|Site), 
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit), type=3) #none

shapiro.test(resid(lm(POM_C ~ AM_hyph_per + AM_arb_per + AM_ves_ct, data = SOM_13C_T0))) #normal
summary(aov(POM_C ~ AM_hyph_per + AM_arb_per + AM_ves_ct, data = SOM_13C_T0)) #hyphae, vesicles almost

Anova(lmer(POM_C ~ AM_hyph_per + AM_arb_per + AM_ves_ct + (1|Site), 
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit), type=3) #vesicles

AM_POM_C_lme <- lmer(POM_C ~ AM_hyph_per + AM_arb_per + AM_ves_ct + (1|Site), 
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit)
summary(AM_POM_C_lme) #random accounts for most variance
shapiro.test(resid(AM_POM_C_lme)) #normal
check_collinearity(AM_POM_C_lme) #no collinearity
performance::r2(AM_POM_C_lme, multivariate=TRUE) #calculate R-squared for model fit
MuMIn::AICc(AM_POM_C_lme, k=2) #calculate AICc for for model comparison with small sample sizes

shapiro.test(resid(lm(POM_N ~ AM_hyph_per + AM_arb_per + AM_ves_ct, data = SOM_13C_T0))) #normal
summary(aov(POM_N ~ AM_hyph_per + AM_arb_per + AM_ves_ct, data = SOM_13C_T0)) #hyphae

Anova(lmer(POM_N ~ AM_hyph_per + AM_arb_per + AM_ves_ct + (1|Site), 
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit), type=3) #hyphae

AM_POM_N_lme <- lmer(POM_N ~ AM_hyph_per + AM_arb_per + AM_ves_ct + (1|Site), 
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = SOM_13C_T0, #choose dataframe
    na.action = na.omit)
summary(AM_POM_N_lme) #
shapiro.test(resid(AM_POM_N_lme)) #normal
check_collinearity(AM_POM_N_lme) #no collinearity
performance::r2(AM_POM_N_lme, multivariate=TRUE) #calculate R-squared for model fit
MuMIn::AICc(AM_POM_N_lme, k=2) #calculate AICc for for model comparison with small sample sizes

shapiro.test(resid(lm(MAOM_C ~ AM_hyph_per + AM_arb_per + AM_ves_ct, data = SOM_13C_T0))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(MAOM_C) ~ AM_hyph_per + AM_arb_per + AM_ves_ct, data = SOM_13C_T0))) #normal
summary(aov(log(MAOM_C) ~ AM_hyph_per + AM_arb_per + AM_ves_ct, data = SOM_13C_T0)) #none

```

```{r AM POM-C chart}

#AM POM-C plot
ggplot(data = SOM_13C_T0, aes(x=AM_ves_ct, y=POM_C)) + 
  geom_point(aes(color = factor(Site))) + #add points to plot
  geom_smooth(method="lm", se=TRUE, color=plotline_color, fill=fill_color) +
  scale_color_manual(values = colors) + #define colors to use
  my_theme +
  #ggtitle("Hyphal Presence Across Sites") + #add a graph title
  xlab("AM Vesicles (Count)") + ylab("POM-C (g C/kg soil)") #define/change X and Y labels

ggsave('POM_C_byAMvesicles.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

```

```{r AM POM-N chart}

#Hyphal presence boxplot
ggplot(data = SOM_13C_T0, aes(x=AM_hyph_per, y=POM_N)) + 
  geom_point(aes(color = factor(Site))) + #add points to plot
  geom_smooth(method="lm", se=TRUE, color=plotline_color, fill=fill_color) +
  scale_color_manual(values = colors) + #define colors to use
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) + #change to %
  my_theme +
  #ggtitle("Hyphal Presence Across Sites") + #add a graph title
  xlab("Presence of AM Hyphae (% of Root)") + ylab("POM-N (g N/kg soil)") #define/change X and Y labels

ggsave('POM_N_byAMhyph.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

```

# AM Association by site

```{r AM presence}

#Hyphal presence boxplot
ggplot(data = AMdata, aes(x=Site, y=hyph_percent, fill=Site)) + 
	geom_boxplot() + 
  scale_fill_manual(values = colors) + #define colors to use
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + #change to %
  my_theme +
  #ggtitle("Hyphal Presence Across Sites") + #add a graph title
  xlab("") + ylab("Presence of AM Hyphae (% of Root)") #define/change X and Y labels

ggsave('AM_hyph.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

shapiro.test(resid(lm(hyph_percent ~ Site, data = AMdata))) #check normality of residuals - normal
summary(aov(hyph_percent ~ Site, data = AMdata)) #sig
TukeyHSD(aov(hyph_percent ~ Site, data = AMdata)) #Flashes > Bick 1 & 2, Ambers > Bick 1


```

```{r AM vesicles}

#Vesicles boxplot
ggplot(data = AMdata, aes(x=Site, y=Vesicle, fill=Site)) + 
	geom_boxplot() + 
  scale_fill_manual(values = colors) + #define colors to use
  #scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + #change to %
  my_theme +
  #ggtitle("Hyphal Presence Across Sites") + #add a graph title
  xlab("") + ylab("AM Vesicles (Count)") #define/change X and Y labels

ggsave('AM_ves.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

shapiro.test(resid(lm(Vesicle ~ Site, data = AMdata))) #check normality of residuals - normal
summary(aov(Vesicle ~ Site, data = AMdata)) #sig
TukeyHSD(aov(Vesicle ~ Site, data = AMdata)) #Flashes > Bick 1 & 2, Ambers > Bick 1


```

# Plant Communities

```{r plant groups}

#order variables
plant_harvest$functional_group <- factor(plant_harvest$functional_group, 
                                        levels = c("grass", "forb", "legume", "shrub"))

#plot plant communities
plant_harvest %>%
  filter(functional_group != "moss") %>% #remove mosses
  filter(functional_group != "unknown") %>% #remove unknown
ggplot(aes(x=site, y=shoot, fill=functional_group)) + 
	geom_boxplot() + 
  geom_jitter(position=position_jitterdodge(jitter.width=0.2, dodge.width=0.75), #add points with jitter
              alpha=0.5, size=0.8, color="black") + #point appearance
  scale_fill_manual(values = colors) + #define colors to use
  #scale_color_manual(values = colors) + #define colors to use
  my_theme +
  #ggtitle("Plant Shoot Biomass Across Sites") + #add a graph title
  xlab("") + ylab("Shoot Biomass (g)") #define/change X and Y labels

#save figure
ggsave('plant_groups.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

shapiro.test(resid(lm(shoot ~ functional_group * site, data = plant_harvest[ #check normality of residuals - not normal
                          !is.na(plant_harvest$shoot), c(1,5,6)]))) #subset only columns of interest, remove na
shapiro.test(resid(lm(log(shoot) ~ functional_group * site, data = plant_harvest[ #normal
                          !is.na(plant_harvest$shoot), c(1,5,6)]))) #subset only columns of interest, remove na
summary(aov(log(shoot) ~ functional_group * site, data = plant_harvest[ 
                          !is.na(plant_harvest$shoot), c(1,5,6)])) #sig for all terms
TukeyHSD(aov(log(shoot) ~ functional_group * site, data = plant_harvest[ #normal
                          !is.na(plant_harvest$shoot), c(1,5,6)])) 
#grass:Bickershaw 2-forb:Bickershaw 1       0.0051386
#legume:Amberswood-grass:Bickershaw 2       0.0241997


```

```{r plant species richness}

#box plots of plant biodiversity metrics
ggplot(diversity, aes(x=site, y=species_richness, fill=site)) + 
  geom_boxplot() +
  scale_fill_manual(values=colors9) + #define colors to use
  scale_y_continuous(breaks = pretty) +
  #ggtitle("Species richness per site") + #add a graph title
  xlab("") + ylab("Species Richness (#)") + #define/change X and Y labels
  #scale_x_discrete(labels = "", "", "", "") + #remove x axis labels
  my_theme 
  #theme(legend.position='bottom') #remove legend

#save figure
ggsave('plant_species.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

#stats
shapiro.test(resid(lm(species_richness ~ site, data = diversity))) #check normality of residuals - normal
summary(aov(species_richness ~ site, data = diversity)) #sig
TukeyHSD(aov(species_richness ~ site, data = diversity)) #Ambers > Bick 1 & 2

```

```{r plant shannon}

#box plots of plant biodiversity metrics
ggplot(diversity, aes(x=site, y=shannon, fill=site)) + 
  geom_boxplot() +
  scale_fill_manual(values=colors9) + #define colors to use
  scale_y_continuous(breaks = pretty) +
  #ggtitle("Species richness per site") + #add a graph title
  xlab("") + ylab("Shannon Diversity Index") + #define/change X and Y labels
  #scale_x_discrete(labels = "", "", "", "") + #remove x axis labels
  my_theme 
  #theme(legend.position='bottom') #remove legend

#save figure
ggsave('plant_shannon.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

#stats
shapiro.test(resid(lm(shannon ~ site, data = diversity))) #check normality of residuals - normal
summary(aov(shannon ~ site, data = diversity)) #not sig

```

```{r plant simpson}

#box plots of plant biodiversity metrics
ggplot(diversity, aes(x=site, y=simpson, fill=site)) + 
  geom_boxplot() +
  scale_fill_manual(values=colors9) + #define colors to use
  scale_y_continuous(breaks = pretty) +
  #ggtitle("Species richness per site") + #add a graph title
  xlab("") + ylab("Simpson Diversity Index") + #define/change X and Y labels
  #scale_x_discrete(labels = "", "", "", "") + #remove x axis labels
  my_theme 
  #theme(legend.position='bottom') #remove legend

#save figure
ggsave('plant_simpson.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

#stats
shapiro.test(resid(lm(simpson ~ site, data = diversity))) #check normality of residuals - normal
summary(aov(simpson ~ site, data = diversity)) #sig
TukeyHSD(aov(simpson ~ site, data = diversity)) #Flashes > Bick 1

```


# Isotope and Incubation Results

## Plant 13C allocation
```{r plant 13C shoots}

#bar plot of plant excess 13C with error bars - shoot
summary_plant13C %>%
  filter(sample_type == "shoot") %>%
ggplot(aes(x=site, y=excess13C_mg_mean, fill=site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  scale_y_continuous(breaks = pretty) +
  geom_errorbar(aes(ymin=excess13C_mg_mean-excess13C_mg_std.error, ymax=excess13C_mg_mean+excess13C_mg_std.error), 
                color = "grey10", width = 0.7, position = "dodge") + 
  #ggtitle("") + #add a graph title
  xlab("") + ylab(expression(""^13*"C in Shoots (mg/pot)")) + #define/change X and Y labels
  my_theme +
  theme(legend.position = "none")

#save figure
ggsave('excess13C_shoots.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications


```

```{r plant 13C roots}

#bar plot of plant excess 13C with error bars - shoot

summary_plant13C %>%
  filter(sample_type == "root") %>%
ggplot(aes(x=site, y=excess13C_mg_mean, fill=site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  #scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, by = 50)) + #set y axis ticks
  geom_errorbar(aes(ymin=excess13C_mg_mean-excess13C_mg_std.error, ymax=excess13C_mg_mean+excess13C_mg_std.error), 
                color = "grey10", width = 0.7, position = "dodge") + 
  #ggtitle("") + #add a graph title
  xlab("") + ylab(expression(""^13*"C in Roots (mg/pot)")) + #define/change X and Y labels
  my_theme +
  theme(legend.position = "none")

#save figure
ggsave('excess13C_roots.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications


```


## Soil isotopes

### Time 0

```{r soil 13C}

#test differences in pre-label soils (natural abundance) 13C
shapiro.test(resid(lm(d13C_MS_pre ~ Site, data = soil_13C_calcs))) #check normality of residuals - normal
summary(aov(d13C_MS_pre ~ Site, data = soil_13C_calcs)) #not sig

#test differences in post-label soils 13C
shapiro.test(resid(lm(T0_d13C_m ~ Site, data = soil_13C_calcs))) #check normality of residuals - normal
summary(aov(T0_d13C_m ~ Site, data = soil_13C_calcs)) #not sig

#test differences in excess 13C in pots at T0
shapiro.test(resid(lm(SOM_pot13C_mg ~ Site, data = SOM_13C_T0))) #check normality of residuals - normal
summary(aov(SOM_pot13C_mg ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(SOM_pot13C_mg ~ Site, data = SOM_13C_T0)) #Flashes > Bick 1, Ambers

#plot excess 13C in pots
summary_SOM_13C %>%
  filter(Fraction == "SOM" & Incubation_T == 0) %>%
ggplot(aes(x=Site, y=pot_13C_mg_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=pot_13C_mg_mean-pot_13C_mg_std.error, ymax=pot_13C_mg_mean+pot_13C_mg_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  #scale_y_continuous(labels = scales::number_format(scale=1*1000, accuracy=1), breaks = pretty) + #change y axis to mg
  #ggtitle("Rocks and Debris per site") + #add a graph title
  xlab("") + ylab(expression(""^13*"C transferred to soil (mg/pot)")) + #define/change X and Y labels
  my_theme +
  theme(legend.position = "none")

#save figure
ggsave('excess13C_soil_050126.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```

```{r soil 13C in POM}

#bar plot of POM 13C with error bars
summary_SOM_13C %>%
  filter(Fraction == "POM" & Incubation_T == 0) %>%
ggplot(aes(x=Site, y=pot_13C_mg_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  scale_y_continuous(breaks = pretty) + #set y axis ticks
  geom_errorbar(aes(ymin=pot_13C_mg_mean-pot_13C_mg_std.error, ymax=pot_13C_mg_mean+pot_13C_mg_std.error), 
                color = "grey10", width = 0.7, position = "dodge") + 
  #ggtitle("") + #add a graph title
  xlab("") + ylab(expression(""^13*"C in POM (mg/pot)")) + #define/change X and Y labels
  my_theme +
  theme(legend.position = "none")

#save figure
ggsave('excess13C_POM_050126.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

shapiro.test(resid(lm(POM_pot13C_mg ~ Site, data = SOM_13C_T0))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(POM_pot13C_mg) ~ Site, data = SOM_13C_T0))) #normal
summary(aov(log(POM_pot13C_mg) ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(log(POM_pot13C_mg) ~ Site, data = SOM_13C_T0)) #Flashes > Bick 1, Ambers

```

```{r soil 13C in MAOM}

#bar plot of POM 13C with error bars
summary_SOM_13C %>%
  filter(Fraction == "MAOM" & Incubation_T == 0) %>%
ggplot(aes(x=Site, y=pot_13C_mg_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  #scale_y_continuous(limits = c(0, 40), breaks = seq(0, 40, by = 10)) + #set y axis ticks
  geom_errorbar(aes(ymin=pot_13C_mg_mean-pot_13C_mg_std.error, ymax=pot_13C_mg_mean+pot_13C_mg_std.error), 
                color = "grey10", width = 0.7, position = "dodge") + 
  #ggtitle("") + #add a graph title
  xlab("") + ylab(expression(""^13*"C in MAOM (mg/pot)")) + #define/change X and Y labels
  my_theme +
  theme(legend.position = "none")

#save figure
ggsave('excess13C_MAOM_050126.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

shapiro.test(resid(lm(MAOM_pot13C_mg ~ Site, data = SOM_13C_T0))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(MAOM_pot13C_mg) ~ Site, data = SOM_13C_T0))) #normal
summary(aov(log(MAOM_pot13C_mg) ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(log(MAOM_pot13C_mg) ~ Site, data = SOM_13C_T0)) #Flashes > Bick 1, Ambers

```

```{r fMAOM-13C - Barker}

#fMAOM C box chart
ggplot(data = SOM_13C_T0, aes(x=Site, y=fMAOM_13C, fill=Site)) + 
	geom_boxplot() +
  scale_fill_manual(values=colors) + #define colors to use
  #ggtitle(expression("fMAOM "^13*"C per site")) + #add a graph title
  xlab("") + 
  ylab(expression("Portion of "^13*"C stored in MAOM")) + #define/change X and Y labels
  scale_y_continuous(labels = scales::percent) + #change y axis to percent
  my_theme

#save figure
ggsave('fMAOM_13C_KB_050126.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

shapiro.test(resid(lm(fMAOM_13C ~ Site, data = SOM_13C_T0))) #check normality of residuals - normal
summary(aov(fMAOM_13C ~ Site, data = SOM_13C_T0)) #almost but not

```

```{r soil 15N}

shapiro.test(resid(lm(SOM_15N ~ Site, data = SOM_13C_T0))) #check normality of residuals - normal
summary(aov(SOM_15N ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(SOM_15N ~ Site, data = SOM_13C_T0)) #Bick 1 < all, Bick 2 < Amber & Flashes

shapiro.test(resid(lm(POM_15N ~ Site, data = SOM_13C_T0))) #check normality of residuals - normal
summary(aov(POM_15N ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(POM_15N ~ Site, data = SOM_13C_T0)) #Bick 1 < Amber & Flashes, Bick 2 < Amber & Flashes

shapiro.test(resid(lm(MAOM_15N ~ Site, data = SOM_13C_T0))) #check normality of residuals - not normal
shapiro.test(resid(lm(log(MAOM_15N) ~ Site, data = SOM_13C_T0))) #normal
summary(aov(log(MAOM_15N) ~ Site, data = SOM_13C_T0)) #sig
TukeyHSD(aov(log(MAOM_15N) ~ Site, data = SOM_13C_T0)) #Bick 1 < Amber & Flashes, Bick 2 < Amber & Flashes


```

### Time 6 Months
```{r soil 13C at 6 mos}

#bar plot of POM 13C with error bars
summary_SOM_13C %>%
  filter(Fraction == "SOM" & Incubation_T == 6) %>%
ggplot(aes(x=Site, y=pot_13C_mg_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  scale_y_continuous(breaks = pretty) + #set y axis ticks
  geom_errorbar(aes(ymin=pot_13C_mg_mean-pot_13C_mg_std.error, ymax=pot_13C_mg_mean+pot_13C_mg_std.error), 
                color = "grey10", width = 0.7, position = "dodge") + 
  #ggtitle("") + #add a graph title
  xlab("") + ylab(expression(""^13*"C in SOM (mg/pot)")) + #define/change X and Y labels
  my_theme +
  theme(legend.position = "none")

#save figure
ggsave('excess13C_6mos.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

shapiro.test(resid(lm(pot_13C_mg ~ Site, data = SOM_T6[-17,]))) #check normality of residuals - normal
summary(aov(pot_13C_mg ~ Site, data = SOM_T6[-17,])) #sig
TukeyHSD(aov(pot_13C_mg ~ Site, data = SOM_T6[-17,])) #Flashes > all

```


## Respiration Data

### 12 C

#### In ppm

```{r resp overall 12}

ggplot(data = resp_full, aes(x=Site, y=CO2_ppm, fill=Site)) + 
	geom_boxplot() + 
  scale_fill_manual(values = colors) + #define colors to use
  my_theme +
  ggtitle("12CO2 Respiration") + #add a graph title
  xlab("") + ylab(expression(""^12*"CO"^2*" (ppm)")) #define/change X and Y labels

ggsave('CO2_overall.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

#ANOVA and Tukey
summary(aov(CO2_ppm ~ factor(Site), data = resp_min)) #almost
summary(aov(CO2_ppm ~ factor(Site), data = resp_full)) #very
TukeyHSD(aov(CO2_ppm ~ factor(Site), data = resp_full)) #Ambers > all, Flashes > Bick 1 & 2

```

```{r resp 12C by week}

#time series boxplots of 12C
ggplot(data = resp_full, aes(x=Week, y=CO2_ppm, group=interaction(Week,Site))) + 
  geom_boxplot(aes(fill=Site), position = position_dodge()) +
  #geom_dotplot(aes(fill=Site, stroke=TRUE), binaxis = "y", dotsize = 0.5, position = position_dodge(), stackdir = "center", binpositions = "bygroup") +
  scale_fill_manual(values = colors) + #define colors to use
  scale_color_manual(values = colors) +
  my_theme +
  scale_y_continuous(name=expression(""^12*"CO"^2*" (ppm)"), breaks = pretty(c(600,8000), n = 8)) + #set limits and change number of tick marks on y axis
  scale_x_continuous(name ="Incubation Time (Week)", breaks = c(0,1)) #set ticks and label for x axis

ggsave('12C_wks.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 7, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications


```

#### In mass of C

```{r C resp mass overall}

multiply1000 <- function(x) {
  return(x * 1000)
}

ggplot(data = resp_full, aes(x=Site, y=mass_C, fill=Site)) + 
	geom_boxplot() + 
  scale_fill_manual(values = colors) + #define colors to use
  my_theme +
  #ggtitle("12CO2 Respiration") + #add a graph title
  scale_y_continuous(name="Total Carbon Respired (g C)", breaks = pretty(c(0,0.2), n=6)) +
  #scale_x_discrete(limits = order_sites) 
  xlab("") + ylab("Total Carbon Respired (mg C)") #define/change X and Y labels

ggsave('CO2_overall_gC.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

#ANOVA and Tukey
summary(aov(mass_C ~ factor(Site), data = resp_full)) #very
TukeyHSD(aov(mass_C ~ factor(Site), data = resp_full)) #Ambers > all, Flashes > Bick 1 & 2

```

```{r resp mass 12C by week}

#time series boxplots of 12C
ggplot(data = resp_full, aes(x=Week, y=mass_C, group=interaction(Week,Site))) + 
  geom_boxplot(aes(fill=Site), position = position_dodge()) +
  #geom_dotplot(aes(fill=Site, stroke=TRUE), binaxis = "y", dotsize = 0.5, position = position_dodge(), stackdir = "center", binpositions = "bygroup") +
  scale_fill_manual(values = colors) + #define colors to use
  scale_color_manual(values = colors) +
  my_theme +
  scale_y_continuous(name="Total Carbon Respired (g C)", breaks = pretty(c(0,0.2), n=6)) + #set limits and change number of tick marks on y axis
  scale_x_continuous(name ="Incubation Time (Week)", breaks = c(0,1)) #set ticks and label for x axis

ggsave('gC_wks.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 7, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications


```



### 13 C

#### In per mille

```{r resp overall 13}

ggplot(data = resp_full, aes(x=Site, y=d13CO2_permille, fill=Site)) + 
	geom_boxplot() + 
  scale_fill_manual(values = colors) + #define colors to use
  my_theme +
  ggtitle("13CO2 Respiration") + #add a graph title
  xlab("") + ylab(expression("Delta "^13*"CO"^2*" (per mille)")) + #define/change X and Y labels
  theme(legend.position='none') #remove legend

ggsave('13CO2_overall.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

#ANOVA and Tukey
summary(aov(d13CO2_permille ~ factor(Site), data = resp_min)) #sig
summary(aov(d13CO2_permille ~ factor(Site), data = resp_full)) #very sig
TukeyHSD(aov(d13CO2_permille ~ factor(Site), data = resp_min)) #23 vs. 15 0.00058, vs. unmanaged almost, 15 vs. 4 0.033
TukeyHSD(aov(d13CO2_permille ~ factor(Site), data = resp_full)) #all sig except Flashes vs. Bick 2

```

```{r resp 13C by week}

#time series boxplots of 12C
ggplot(data = resp_full, aes(x=Week, y=d13CO2_permille, group=interaction(Week,Site))) + 
  geom_boxplot(aes(fill=Site), position = position_dodge()) +
  #geom_dotplot(aes(fill=Site, stroke=TRUE), binaxis = "y", dotsize = 0.5, position = position_dodge(), stackdir = "center", binpositions = "bygroup") +
  scale_fill_manual(values = colors) + #define colors to use
  scale_color_manual(values = colors) +
  my_theme +
  scale_y_continuous(name=expression("Delta "^13*"CO"^2*" (per mille)"), breaks = pretty(c(0,1000), n = 8)) + #set limits and change number of tick marks on y axis
  scale_x_continuous(name ="Weeks of Incubation", breaks = c(0,1,2,3,4)) #set ticks and label for x axis

ggsave('13C_wks.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 7, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications


```

#### In percent excess 13C

```{r resp overall 13 excess}

ggplot(data = resp_full, aes(x=Site, y=atom_excess13C, fill=Site)) + 
	geom_boxplot() + 
  scale_fill_manual(values = colors) + #define colors to use
  my_theme +
  #ggtitle("13CO2 Respiration") + #add a graph title
  scale_y_continuous(name = expression("Excess "^13*"C Respiration (atom %)"), labels = scales::percent) + #change y axis to percent, label
  xlab("") + #define/change X label
  theme(legend.position='none') #remove legend

ggsave('13CO2_overall_atom.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

#ANOVA and Tukey
summary(aov(atom_excess13C ~ factor(Site), data = resp_full)) #very sig
TukeyHSD(aov(atom_excess13C ~ factor(Site), data = resp_full)) #all sig except Flashes vs. Bick 2

```

```{r resp excess 13C by week}

#time series boxplots of 12C
ggplot(data = resp_full, aes(x=Week, y=atom_excess13C, group=interaction(Week,Site))) + 
  geom_boxplot(aes(fill=Site), position = position_dodge()) +
  #geom_dotplot(aes(fill=Site, stroke=TRUE), binaxis = "y", dotsize = 0.5, position = position_dodge(), stackdir = "center", binpositions = "bygroup") +
  scale_fill_manual(values = colors) + #define colors to use
  scale_color_manual(values = colors) +
  my_theme +
  scale_y_continuous(name = expression("Excess "^13*"C Respiration (atom %)"), labels = scales::percent) + #change y axis to percent, label
  scale_x_continuous(name ="Weeks of Incubation", breaks = c(0,1,2,3,4)) #set ticks and label for x axis

ggsave('excess13C_wks.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 7, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications


```

#### In mass of excess 13C

```{r resp overall 13C mass}


ggplot(data = resp_full, aes(x=Site, y=mass_excess13C, fill=Site)) + 
	geom_boxplot() + 
  scale_fill_manual(values = colors) + #define colors to use
  my_theme +
  #ggtitle("13CO2 Respiration") + #add a graph title
  scale_y_continuous(name = expression("Total "^13*"C Respired (mg "^13*"C)"), labels = multiply1000) + #change y axis, label
  xlab("") + #define/change X label
  theme(legend.position='none') #remove legend

ggsave('13C_overall_mass.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 6, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

#ANOVA and Tukey
summary(aov(mass_excess13C ~ factor(Site), data = resp_full)) #very sig
TukeyHSD(aov(mass_excess13C ~ factor(Site), data = resp_full)) #all sig except Flashes vs. Bick 2

```

```{r resp 13C mass by week}

#time series boxplots
ggplot(data = resp_full, aes(x=Week, y=mass_excess13C, group=interaction(Week,Site))) + 
  geom_boxplot(aes(fill=Site), position = position_dodge()) +
  #geom_dotplot(aes(fill=Site, stroke=TRUE), binaxis = "y", dotsize = 0.5, position = position_dodge(), stackdir = "center", binpositions = "bygroup") +
  scale_fill_manual(values = colors) + #define colors to use
  scale_color_manual(values = colors) +
  my_theme +
  scale_y_continuous(name = expression("Total "^13*"C Respired (mg "^13*"C)"), labels = multiply1000) + #change y axis, label
  scale_x_continuous(name ="Weeks of Incubation", breaks = c(0,1,2,3,4)) #set ticks and label for x axis

ggsave('excess13C_wks_mass.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 7, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications


```


## Root:shoot ratio
```{r rs, warning=FALSE}

#root:shoot ratio box plot
ggplot(data = plant_harvest_est[-c(118,153),], #species data minus major outliers
       aes(x=site, y=root_shoot, fill=site)) + 
	geom_boxplot() +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("Root:Shoot ratio per site") + #add a graph title
  xlab("Site") + ylab("Root:Shoot ratio") + #define/change X and Y labels
  my_theme

```

## Biomass

### Beginning combined roots
```{r beg root bio, warning=FALSE}

#mass bar chart with error bars
ggplot(data = summary_endbio, aes(x=Site, y=est_beg_drymass_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=est_beg_drymass_mean-est_beg_drymass_std.error, ymax=est_beg_drymass_mean+est_beg_drymass_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Beginning biomass per site") + #add a graph title
  xlab("Site") + ylab("Root dry mass (g)") + #define/change X and Y labels
  my_theme

ggsave('beg_roots.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 7, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

#ANOVA and Tukey
summary(aov(est_beg_drymass ~ factor(Site), data = endbiomass)) #sig
TukeyHSD(aov(est_beg_drymass ~ factor(Site), data = endbiomass)) #Flashes < Amberswood & Bick 2

```

### Ending combined roots
```{r end root bio, warning=FALSE}

#mass bar chart with error bars
ggplot(data = summary_endbio, aes(x=Site, y=end_drymass_clean_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=end_drymass_clean_mean-end_drymass_clean_std.error, ymax=end_drymass_clean_mean+end_drymass_clean_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  ggtitle("Ending biomass per site") + #add a graph title
  xlab("Site") + ylab("Root dry mass (g)") + #define/change X and Y labels
  my_theme

ggsave('end_roots.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
       path = "Figures", #save to folder
       width = 7, height = 4, units = "in", #set dimensions
       dpi = 300, #set resolution
       bg=NULL) #background from plot specifications

#ANOVA and Tukey
summary(aov(end_drymass_clean ~ factor(Site), data = endbiomass)) #sig
TukeyHSD(aov(end_drymass_clean ~ factor(Site), data = endbiomass)) #Flashes < Bick 1, almost Bick 2

```

### Decomposition

```{r decomp, warning=FALSE}

#decomposition percentage box plot
endbiomass[-1,] %>% #remove G37 because of biomass est error
ggplot(aes(x=Site, y=decomp_per_adj)) + 
	geom_boxplot(aes(fill=Site)) +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("Decomposition per Site") + #add a graph title
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + #change to %
  xlab("") + ylab("Decomposition (%)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('decomp.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications


```

```{r decomp plot restoration time}

#decomposition lm plot vs restoration time
ggplot(data = endbiomass[-1,], aes(x=years, y=decomp_per_adj)) + 
	geom_point(aes(color=Management), size=2, alpha=0.5) +
  scale_fill_manual(values=colors) + #define colors to use
  geom_smooth(method="lm", se=TRUE, color=plotline_color, fill=fill_color) +
  scale_x_continuous(breaks=c(0,4,15,23), expand = c(0.1,0.1)) + #set limits and change tick marks on x axis
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + #change to %
  ggtitle("Decomposition by Restoration Time") + #add a graph title
  xlab("Management Time") + ylab("Decomposition (%)") + #define/change X and Y labels
  my_theme

```

```{r decomp plot pH}

#decomposition lm plot vs restoration time
ggplot(data = endbiomass[-1,], aes(x=pH, y=decomp_per_adj)) + 
	geom_point(aes(color=Site), size=2) +
  scale_color_manual(values=colors) + #define colors to use
  geom_smooth(method="lm", se=TRUE, color=plotline_color, fill=fill_color) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + #change to %
  ggtitle("Decomposition by pH") + #add a graph title
  xlab("pH") + ylab("Decomposition (%)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('decomp_pH.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

```


```{r decomp stats}

#ANOVA and Tukey - Site comparison
summary(aov(decomp_per_adj ~ factor(Site), data = endbiomass[-1,])) #sig
TukeyHSD(aov(decomp_per_adj ~ factor(Site), data = endbiomass[-1,])) #Flashes > Bick 1 & 2, Amberswood > Bick 1
kruskal.test(decomp_per_adj ~ factor(Site), data = endbiomass[-1,]) #Kruskal-Wallis rank sum test also sig
dunn.test(endbiomass[-1,]$decomp_per_adj, endbiomass[-1,]$Site, method="bonferroni", kw=TRUE) #KW P=0.01, Flashes > Bick 1 & 2 only

#AM association
shapiro.test(resid(lm(decomp_per_adj ~ AM_hyph_per + AM_arb_per + AM_ves_ct + AM_absent, data = endbiomass))) #check normality of residuals - normal
summary(aov(decomp_per_adj ~ AM_hyph_per + AM_arb_per + AM_ves_ct + AM_absent, data = endbiomass)) #none

#pH
shapiro.test(resid(lm(decomp_per_adj ~ pH, data = endbiomass))) #check normality of residuals - normal
summary(aov(decomp_per_adj ~ pH, data = endbiomass)) #very sig

Anova(lmer(decomp_per_adj ~ pH + (1|Site), 
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = endbiomass, #choose dataframe
    na.action = na.omit), type=3) #pH

decomp_pH_lme <- lmer(decomp_per_adj ~ pH + (1|Site), 
    REML = TRUE, #restricted maximum likelihood (REML) for model comparison 
    data = endbiomass, #choose dataframe
    na.action = na.omit)
summary(decomp_pH_lme) #random accounts for most variance
shapiro.test(resid(decomp_pH_lme)) #normal
performance::r2(decomp_pH_lme, multivariate=TRUE) #calculate R-squared for model fit
MuMIn::AICc(decomp_pH_lme, k=2) #calculate AICc for for model comparison with small sample sizes

```


## Rocks and debris

```{r rocks, warning=FALSE}

#mass bar chart with error bars
ggplot(data = summary_endbio, aes(x=Site, y=rocks_debris_mean, fill=Site)) + 
	geom_bar(stat = "identity", width = 0.7, position = "dodge") +
  scale_fill_manual(values=colors) + #define colors to use
  geom_errorbar(aes(ymin=rocks_debris_mean-rocks_debris_std.error, ymax=rocks_debris_mean+rocks_debris_std.error),
  color = "grey10", width = 0.7, position = "dodge") + 
  #ggtitle("Rocks and Debris per site") + #add a graph title
  xlab("") + ylab("Mass of Coal Debris in 20 cm (g)") + #define/change X and Y labels
  my_theme

#save figure
ggsave('debris.svg', #save plot name and type - can be "eps", "pdf", "jpeg", "tiff", "png", "bmp", "svg"
	path = "Figures", #save to folder within working directory
	width = 6, height = 4, units = "in", #set dimensions for figure
	dpi = 300, #set resolution
	bg=NULL) #background from plot specifications

#ANOVA and Tukey
rmass_aov <- aov(rocks_debris ~ factor(Management), data = endbiomass)
summary(rmass_aov) #P=2.3e-6
TukeyHSD(rmass_aov) #Unmanaged-23 yr P=0.0000012, Unmanaged-15 yr P=0.0057, 4 yr-23 yr P=0.00013, 15 yr-23 yr P=0.0013

```

## Worms

```{r worms, warning=FALSE}

#worms box plot
ggplot(data = endbiomass, aes(x=Management, y=worms, fill=Management)) + 
	geom_boxplot() +
  scale_fill_manual(values=colors) + #define colors to use
  ggtitle("Worms per site") + #add a graph title
  xlab("Management") + ylab("Number of worms") + #define/change X and Y labels
  my_theme

```
